---
title: "Design document for ocRctPredprobDist"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

Calculate operating characteristics for RCT against SOC,
using the predictive probability method with beta priors

We emulate a randomized-controlled trial setting where at any given sample size,
there exists the number of patients enrolled in either standard of care (SOC) or control arm, and
a treatment or experimental arm. The allocation of patients will depend on the
randomization ratio set by the user and is rounded to the next higher integer.
Therefore the sequence of patients is determined from the start, such that the number of
patients in both arms is constant across trial simulations, however the number of patients
within the control and treatment arm is determined by the randomisation ratio.
Interim looks are for sample sizes below that of the final sample size.

Final looks are only performed at the maximum sample size.

At each interim or final look, a futility or efficacy or both can be performed.

The rules for Stop, Go and Gray Zone (where applicable) are the same as in [ocPredprobDist()] where
the only difference is to emulate a randomized-controlled trial setting.

The returned value is a list with the following elements:
 - `oc`: matrix with operating characteristics with the following details:
    - `ExpectedN`: expected number of patients in the trials in both treatment and SOC group
    - `ExpectedNactive` : the mean of the number of patients in treatment arm
    - `ExpectedNcontrol`: the mean of the number of patients in control arm
    - `PrStopEarly`: probability to stop the trial early (before reaching the
                    maximum sample size)
    - `PrEarlyEff`: probability of Early Go decision
    - `PrEarlyFut`: probability of Early Stop decision
    - `PrEfficacy`: probability of Go decision
    - `PrFutility`: probability of Stop decision
    - `PrGrayZone`: probability of Evaluate or "Gray Zone" decision (between Go and Stop)
 - `Decision` : numeric of results with `TRUE` as Go, `FALSE` as Stop and `NA` as Evaluate decision.
 - `SampleSize` : numeric of sample sizes from `nnE` or `nnF` or both.
 - `wiggled_nnE` : user input for `nnE` with random distance applied.
 - `wiggled_nnF` : user input for `nnF` with random distance applied.
 - `wiggled_dist` : magnitude of random distance applied in order of input looks.
 - `params` : all user input arguments.

## Input parameters
```{r}
devtools::load_all()
input <- list(
  xS = 0,
  nS = 0,
  truep = 0.4,
  nn = c(10, 20, 30),
  nnE = c(10, 20, 30),
  pE = 0.4,
  pS = 0.3,
  deltaE = 0.2,
  deltaF = 0.1,
  relativeDelta = FALSE,
  tT = 0.5,
  tL = 0.8,
  tU = 0.8,
  phiU = 0.5,
  phiL = 0.3,
  parE = c(a = 1, b = 1),
  parS = c(a = 1, b = 1),
  randRatio = 1,
  ns = 10000,
  nr = FALSE,
  d = NULL,
  nnF = c(10, 20, 30),
  truep = 0.4,
  sim = 5,
  Nmax = 15,
  NmaxControl = 15,
  decision <- rep(NA, 5),
  weights = 1,
  weightS = 1
)
n <- nActive <- nControl <- integer(5) # initiate n
ns <- as.integer(5) # sim
nL <- length(input$nn) # the max of nn or we can do Nmax
Nstart <- input$nn[1]
Nmax <- input$nn[nL]
```

## Example of vintage function

```{r}
ocRctPredprobDist(
  nn = input$nn,
  pE = input$pE,
  pS = input$pS,
  delta = input$deltaE,
  deltaFu = input$deltaF,
  relativeDelta = input$relativeDelta,
  tT = 0.6,
  phiL = 0.2,
  phiU = 0.6,
  parE = input$parE,
  parS = input$parS,
  randRatio = 1,
  ns = input$sim,
  nr = input$nr,
  d = input$d,
  nnF = input$nnF
)
```

# h_decision_one_rct_predprobDist
```{r}
h_decision_one_rct_predprobDist <- function(
    nnr,
    nnE,
    nnF,
    xS,
    nS,
    pE,
    pS,
    parE,
    parS,
    tT,
    phiU,
    phiL,
    deltaE,
    deltaF,
    weights,
    weightsS,
    relativeDelta = FALSE,
    randRatio = 1) {
  assert_numeric(nnE, lower = 1, sorted = TRUE, any.missing = FALSE)
  assert_numeric(nnF, lower = 1, sorted = TRUE, any.missing = FALSE)
  assert_numeric(nnr, lower = 1, sorted = TRUE)
  assert_numeric(parE, min.len = 2, any.missing = FALSE)
  assert_numeric(parS, min.len = 2, any.missing = FALSE)
  assert_number(tT, lower = 0, upper = 1)
  assert_number(phiU, lower = 0, upper = 1)
  assert_number(phiL, lower = 0, upper = 1)
  assert_number(deltaE, finite = TRUE, upper = 1)
  assert_number(deltaF, finite = TRUE, upper = 1)
  assert_number(weights, lower = 1, finite = TRUE)
  assert_number(weightsS, lower = 1, finite = TRUE)
  assert_flag(relativeDelta)

  Nmax <- max(nnr)
  NmaxControl <- max(nnF)
  index_look <- 1
  size_look <- nnr[index_look]
  all_sizes <- decision <- nActive <- nControl <- NA
  activeProp <- randRatio / (randRatio + 1)

  NmaxActive <- ceiling(activeProp * Nmax)
  NmaxControl <- Nmax - NmaxActive

  isActive <- sample(
    x = rep(c(TRUE, FALSE), c(NmaxActive, NmaxControl)),
    size = Nmax,
    replace = FALSE
  )
  response <- rbinom(Nmax, size = 1, prob = ifelse(isActive, pE, pS))

  ### Decision 1:
  # The criteria for Decision 1 for Interim looks are :
  # Interim GO =  P(successful trial at final) > phiU
  # Interim STOP = P(successful trial at final) < phiL
  while (is.na(decision) && index_look < length(nnr)) {
    xActive <- response[which(isActive[1:size_look])]
    xControl <- response[which(!isActive[1:size_look])]

    size_look <- nnr[index_look]
    if (size_look %in% nnE) {
      interim_qU <- predprobDist(
        x = sum(xActive),
        n = length(xActive),
        xS = sum(xControl),
        nS = length(xControl),
        Nmax = Nmax,
        NmaxControl = NmaxControl,
        delta = deltaE,
        relativeDelta = relativeDelta,
        parE = parE,
        weights = weights,
        parS = parS,
        weightsS = weightsS,
        thetaT = tT
      )$result
      decision <- ifelse(interim_qU > phiU, TRUE, decision)
    }
    if (size_look %in% nnF) {
      interim_qU <- predprobDist(
        x = sum(xActive),
        n = length(xControl),
        xS = sum(xControl),
        nS = length(xControl),
        Nmax = Nmax,
        NmaxControl = NmaxControl,
        delta = deltaF,
        relativeDelta = relativeDelta,
        parE = parE,
        weights = weights,
        parS = parS,
        weightsS = weightsS,
        thetaT = tT
      )$result
      decision <- ifelse(interim_qU < phiL, FALSE, decision)
    }
    index_look <- index_look + 1
    # The criteria for Decision 1 for Final looks are:
    # Final GO = P(RR > p0 + deltaE | data) => tT
    # Final STOP = P(RR > p0 + deltaF | data ) < tT
    if (is.na(decision)) {
      size_look <- nnr[index_look]
      if (size_look %in% nnE) {
        final_eff_qU <- postprobDist(
          x = sum(xActive),
          n = length(xControl),
          xS = sum(xControl),
          nS = length(xControl),
          delta = deltaE,
          relativeDelta = relativeDelta,
          parE = parE,
          weights = weights, # for activeBetamixPost
          parS = parS,
          weightsS = weightsS # for controlBetamixPost
        )
        decision <- ifelse(final_eff_qU >= tT, TRUE, NA)
      }
      if (size_look %in% nnF) {
        final_fu_qU <- postprobDist(
          x = sum(xActive),
          n = length(xControl),
          xS = sum(xControl),
          nS = length(xControl),
          delta = deltaF,
          relativeDelta = relativeDelta,
          parE = parE,
          weights = weights, # for activeBetamixPost
          parS = parS,
          weightsS = weightsS # for controlBetamixPost
        )
        decision <- ifelse(final_fu_qU < tT, FALSE, decision)
      }
    }
  }
  list(
    nActive = length(xActive),
    nControl = length(xControl),
    decision = decision,
    all_sizes = size_look
  )
}
```

## h_decision_one_rct_predprobDist

```{r}
example_dist <- h_get_distance(nn = c(10, 20, 30))
looks <- h_get_looks(dist = example_dist, nnE = c(10, 20, 30), nnF = c(10, 20, 30))
looks_nnrE <- looks$nnrE
looks_nnrF <- looks$nnrF

h_decision_one_rct_predprobDist(
  nnr = c(10, 20, 30),
  nnE = looks$nnrE,
  nnF = looks$nnrF,
  pE = 0.6,
  pS = 0.3,
  parE = c(1, 1),
  parS = c(1, 1),
  tT = 0.7,
  phiU = 0.3,
  phiL = 0.7,
  deltaE = 0.1,
  deltaF = 0.1,
  weights = 1,
  weightsS = 1,
  relativeDelta = FALSE,
  randRatio = 1
)
```

# h_decision_two_rct_predprobDist

```{r}
h_decision_two_rct_predprobDist <- function(
    nnE,
    nnF,
    nnr,
    pE,
    pS,
    parE,
    parS,
    tF,
    tT,
    phiU,
    phiFu,
    deltaE,
    deltaF,
    weights,
    weightsS,
    relativeDelta,
    randRatio) {
  assert_numeric(nnE, lower = 1, sorted = TRUE, any.missing = FALSE)
  assert_numeric(nnF, lower = 1, sorted = TRUE, any.missing = FALSE)
  assert_numeric(nnr, lower = 1, sorted = TRUE, any.missing = FALSE)
  assert_numeric(parE, min.len = 2, any.missing = FALSE)
  assert_numeric(parS, min.len = 2, any.missing = FALSE)
  assert_number(tT, lower = 0, upper = 1)
  assert_number(tF, lower = 0, upper = 1)
  assert_number(phiU, lower = 0, upper = 1)
  assert_number(phiFu, lower = 0, upper = 1)
  assert_number(deltaE, upper = 1)
  assert_number(deltaF, upper = 1)
  assert_number(weights, lower = 1, finite = TRUE)
  assert_number(weightsS, lower = 1, finite = TRUE)
  assert_flag(relativeDelta)

  Nmax <- max(nnr)
  NmaxControl <- max(nnF)
  index_look <- 1
  size_look <- nnr[index_look]
  all_sizes <- decision <- nActive <- nControl <- NA
  activeProp <- randRatio / (randRatio + 1)

  NmaxActive <- ceiling(activeProp * Nmax)
  NmaxControl <- Nmax - NmaxActive

  isActive <- sample(
    x = rep(c(TRUE, FALSE), c(NmaxActive, NmaxControl)),
    size = Nmax,
    replace = FALSE
  )
  response <- rbinom(Nmax, size = 1, prob = ifelse(isActive, pE, pS))

  # The criteria for Decision 2 for Interim looks are :
  # Interim GO : P ( successful at final) > phiU
  # Interim STOP : P ( unsuccessful at final ) > phiFu
  while (is.na(decision) && index_look < length(nnr)) {
    xActive <- response[which(isActive[1:size_look])]
    xControl <- response[which(!isActive[1:size_look])]

    size_look <- nnr[index_look]
    if (size_look %in% nnE) {
      interim_qU <- predprobDist(
        x = sum(xActive),
        n = length(xActive),
        xS = sum(xControl),
        nS = length(xControl),
        Nmax = Nmax,
        NmaxControl = NmaxControl,
        delta = deltaE,
        relativeDelta = relativeDelta,
        parE = parE,
        weights = weights,
        parS = parS,
        weightsS = weightsS,
        thetaT = tT
      )$result
      decision <- ifelse(interim_qU > phiU, FALSE, decision)
    }
    if (size_look %in% nnF) {
      interim_qU <- 1 - predprobDist(
        x = sum(xActive),
        n = length(xActive),
        xS = sum(xControl),
        nS = length(xControl),
        Nmax = Nmax,
        NmaxControl = NmaxControl,
        delta = deltaF,
        relativeDelta = relativeDelta,
        parE = parE,
        weights = weights,
        parS = parS,
        weightsS = weightsS,
        thetaT = tT
      )$result
      decision <- ifelse(interim_qU > phiFu, FALSE, decision)
    }
    index_look <- index_look + 1
  }
  # The criteria for Decision 2 for Futility looks are :
  # Final GO = P( RR > p0 + delta ) > tT
  # Final STOP = P( RR < p1 - delta ) > tF
  if (is.na(decision)) {
    size_look <- nnr[index_look]
    if (size_look %in% nnE) {
      final_eff_qU <- postprobDist(
        x = sum(xActive),
        n = length(xControl),
        xS = sum(xControl),
        nS = length(xControl),
        delta = deltaE,
        relativeDelta = relativeDelta,
        parE = parE,
        parS = parS
      )
    }
    decision <- ifelse(final_eff_qU > tT, TRUE, NA)
  }
  if (size_look %in% nnF) {
    final_fu_qU <- 1 - postprobDist(
      x = sum(xActive),
      n = length(xControl),
      xS = sum(xControl),
      nS = length(xControl),
      delta = deltaF,
      relativeDelta = relativeDelta,
      parE = parE,
      parS = parS
    )
    decision <- ifelse(final_fu_qU > tF, FALSE, decision)
  }
  list(
    decision = decision,
    all_sizes = size_look,
    nActive = length(xActive),
    nControl = length(xControl)
  )
}
```
# h_decision_two_rct_predprobDist example when `relativeDelta = TRUE`
```{r}
h_decision_two_rct_predprobDist(
  nnE = input$nnE,
  nnF = input$nnF,
  nnr = input$nn,
  pE = 0.6,
  pS = 0.3,
  parE = input$parE,
  parS = input$parS,
  tT = input$tT,
  tF = 0.4,
  phiU = input$phiU,
  phiFu = 0.6,
  deltaE = input$deltaE,
  deltaF = input$deltaF,
  weights = input$weights,
  weightsS = input$weightS,
  relativeDelta = TRUE,
  randRatio = 1
)
```
# h_decision_two_rct_predprobDist example when `relativeDelta = FALSE`
```{r}
h_decision_two_rct_predprobDist(
  nnE = input$nnE,
  nnF = input$nnF,
  nnr = input$nn,
  pE = 0.6,
  pS = 0.3,
  parE = input$parE,
  parS = input$parS,
  tT = input$tT,
  tF = 0.4,
  phiU = input$phiU,
  phiFu = 0.6,
  deltaE = input$deltaE,
  deltaF = input$deltaF,
  weights = input$weights,
  weightsS = input$weightS,
  relativeDelta = FALSE,
  randRatio = 1
)
```
# h_get_oc_rct can be used from `ocRctPostprobDist()`
```{r}
h_get_oc_rct <- function(all_sizes, Nmax, nActive, nControl, decision) {
  assert_numeric(nActive, any.missing = FALSE, len = length(all_sizes))
  assert_numeric(nControl, any.missing = FALSE, len = length(all_sizes))
  assert_true(all(nActive + nControl == all_sizes))
  assert_true(Nmax >= max(all_sizes))

  tmp <- h_get_oc(
    all_sizes = all_sizes,
    Nmax = Nmax,
    decision = decision
  )
  tmp$ExpectedNactive <- mean(nActive)
  tmp$ExpectedNcontrol <- mean(nControl)
  tmp
}
```

# Example
```{r}
h_get_oc_rct(
  all_sizes = c(10, 20, 30),
  Nmax = 30,
  nActive = c(4, 10, 15),
  nControl = c(6, 10, 15),
  decision = c(TRUE, FALSE, FALSE)
)
```


# new user facing function

```{r}
new_ocRctPredprobDist <- function(nnE,
                                  pE,
                                  pS,
                                  deltaE,
                                  deltaF,
                                  phiU,
                                  phiL,
                                  relativeDelta = FALSE,
                                  tT,
                                  tF,
                                  parE = c(a = 1, b = 1),
                                  parS = c(a = 1, b = 1),
                                  weights,
                                  weightsS,
                                  randRatio = 1,
                                  sim,
                                  wiggle = FALSE,
                                  nnF = nnE,
                                  decision1 = TRUE) {
  assert_numeric(nnE, min.len = 1, lower = 1, upper = max(nnE), any.missing = FALSE)
  # assert_number(pE, lower = 0, upper = 1)
  # assert_number(pS, lower = 0, upper = 1)
  assert_number(deltaE, upper = 1, finite = TRUE)
  assert_number(deltaF, upper = 1, finite = TRUE)
  assert_flag(relativeDelta)
  # assert_number(tL, lower = 0, upper = 1)
  # assert_number(tU, lower = 0, upper = 1)
  assert_numeric(parE, lower = 0, finite = TRUE, any.missing = FALSE)
  assert_numeric(parS, lower = 0, finite = TRUE, any.missing = FALSE)
  assert_number(sim, lower = 1, finite = TRUE)
  assert_flag(wiggle)
  assert_numeric(nnF, min.len = 0, any.missing = FALSE)

  # some starting initiation ocRctPostprobDist
  if (sim < 50000) {
    warning("Advise to use sim >= 50000 to achieve convergence")
  }
  decision <- all_sizes <- nActive <- nControl <- vector(length = sim)
  nnE <- sort(nnE)
  nnF <- sort(nnF)
  nn <- sort(unique(c(nnF, nnE)))
  nL <- length(nn)
  nn <- sort(unique(c(nnF, nnE)))

  Nstart <- nn[1]
  Nmax <- nn[nL]

  for (k in seq_len(sim)) {
    if (length(nn) != 1 && wiggle) {
      dist <- h_get_distance(nn = nn)
      nnr <- h_get_looks(dist = dist, nnE = nnE, nnF = nnF)
      nnrE <- nnr$nnrE
      nnrF <- nnr$nnrF
    } else {
      nnrE <- nnE
      nnrF <- nnF
    }
    nnr <- unique(c(nnrE, nnrF))
    tmp <- if (decision1) {
      h_decision_one_rct_predprobDist(
        nnE = nnE,
        nnF = nnF,
        nnr = nnr,
        xS = xS,
        nS = nS,
        pE = pE,
        pS = pS,
        parE = parE,
        parS = parS,
        tT = tT,
        phiU = phiU,
        phiL = phiL,
        deltaE = deltaE,
        deltaF = deltaF,
        weights = weights,
        weightsS = weightsS,
        relativeDelta = relativeDelta
      )
    } else {
      tmp <- h_decision_two_rct_predprobDist(
        nnE = nnE,
        nnF = nnF,
        nnr = nnr,
        xS = xS,
        nS = nS,
        pE = pE,
        pS = pS,
        parE = parE,
        parS = parS,
        tT = tT,
        tF = tF,
        phiU = phiU,
        phiFu = phiFu,
        deltaE = deltaE,
        deltaF = deltaF,
        weights = weights,
        weightsS = weightsS,
        relativeDelta = relativeDelta,
        randRatio = randRatio
      )
    }
    decision[k] <- tmp$decision
    all_sizes[k] <- tmp$all_sizes
    nActive[k] <- tmp$nActive
    nControl[k] <- tmp$nControl
  }
  oc <- h_get_oc_rct(
    all_sizes = all_sizes,
    Nmax = Nmax,
    nActive = nActive,
    nControl = nControl,
    decision = decision
  )
  list(
    oc = oc,
    nActive = nActive,
    nControl = nControl,
    ExpectedN = mean(all_sizes),
    ExpectedNactive = mean(nActive),
    ExpectedNcontrol = mean(nControl),
    Decision = decision,
    SampleSize = all_sizes,
    SampleSizeActive = nActive,
    SampleSizeControl = nControl,
    union_nn = nnr,
    wiggled_nnE = nnrE,
    wiggled_nnF = nnrF,
    wiggle_dist = dist,
    params = as.list(match.call(expand.dots = FALSE))
  )
}

# Daniel: runs ok until here, afterwards need to still work on the design, because here you don't develop new_RctPostProbDist but new_ocRctPredprobDist
```


# Example of new_RctPostProbDist when `relativeDelta = TRUE` and `decision1 = TRUE` comparing with vintage function below:

<!--  -->
<!-- Go interim: -->
<!-- P(success at final) > phiU -->
<!-- Go final: -->
<!-- P(P_E > p0 + deltaE) > tT -->

<!-- Stop interim: -->
<!-- P(failure at final) > phiU -->
<!-- Stop final: -->
<!-- P(P_S < p1 + deltaF) > tFu -->

```{r}
set.seed(2000)
new_ocRctPredprobDist(
  nnE = c(10, 20, 30),
  pE = 0.4,
  pS = 0.2,
  deltaE = input$deltaE,
  deltaF = input$deltaF,
  phiU = 0.2,
  phiL = 0.8,
  relativeDelta = TRUE,
  tT = 0.6,
  parE = input$parE,
  parS = c(a = 1, b = 1),
  weights = 1,
  weightsS = 1,
  randRatio = 1,
  sim = 20,
  wiggle = FALSE,
  nnF = c(10, 20, 30),
  decision1 = TRUE
)
```

```{r}
# set.seed(2000)
# ocRctPredprobDist(
#   nn = input$nn,
#   pE = input$pE,
#   pS = input$pS,
#   # Daniel: this does not work
#   deltaE = input$deltaE,
#   deltaF = input$deltaF,
#   delta = input$deltaE,
#   deltaFu = input$deltaF,
#   relativeDelta = TRUE,
#   tL = input$tL,
#   tU = input$tU,
#   parE = input$parE,
#   parS = input$parS,
#   randRatio = 1,
#   ns = 50,
#   nr = FALSE,
#   d = NULL,
#   nnF = input$nnF
# )
```

# Example of new_ocRctPostprobDist when `relativeDelta = FALSE`.

```{r}
set.seed(2000)
new_RctPostprobDist(
  nnE = input$nn,
  pE = input$pE,
  pS = input$pS,
  deltaE = input$deltaE,
  deltaF = input$deltaF,
  relativeDelta = FALSE,
  tL = input$tL,
  tU = input$tU,
  parE = input$parE,
  parS = input$parE,
  randRatio = 1,
  sim = 50,
  wiggle = FALSE,
  nnF = input$nnF
)
```

# Example of ocRctPostprobDist when `relativeDelta = FALSE`.

```{r}
set.seed(1090)
ocRctPostprobDist(
  nn = input$nn,
  pE = input$pE,
  pS = input$pS,
  deltaE = input$deltaE,
  deltaF = input$deltaF,
  relativeDelta = FALSE,
  tL = input$tL,
  tU = input$tU,
  parE = c(a = 1, b = 1),
  parS = c(a = 1, b = 1),
  randRatio = 1,
  ns = 50,
  nr = FALSE,
  d = NULL,
  nnF = input$nnF
)
```


# vintage function

```{r}
ocRctPredprobDist <- function(nn, pE, pS, delta = 0, deltaFu = delta, relativeDelta = FALSE,
                              tT, tFu = 1 - tT,
                              phiL = 1 - phiFu, phiU, phiFu = 1 - phiL,
                              parE = c(a = 1, b = 1), parS = c(a = 1, b = 1),
                              randRatio = 1,
                              ns = 10000, nr = FALSE, d = NULL, nnF = nn) {
  ## checks
  stopifnot(
    is.probability(pE),
    is.probability(pS),
    is.probability(delta),
    is.bool(relativeDelta),
    is.probability(tT),
    is.probability(phiL),
    is.probability(phiU),
    randRatio > 0,
    is.scalar(ns),
    is.bool(nr)
  )

  ## s: decision reject H0 (TRUE) or fail to reject (FALSE)
  ##    during trial if continuing (NA)

  if (phiL + phiFu != 1) {
    warning("Both phiL and phiFu arguments are specified, phiL will be overwrite by 1-phiFu")
  }

  nnE <- sort(nn)
  nnF <- sort(nnF)
  s <- rep(NA, ns)
  n <- nActive <- nControl <- s
  ns <- as.integer(ns)
  nn <- sort(unique(c(nnF, nnE)))
  nL <- length(nn)
  Nstart <- nn[1]
  Nmax <- nn[nL]

  ## proportion of active patients:
  activeProp <- randRatio / (randRatio + 1)

  ## determine number of active and control patients
  NmaxActive <- ceiling(activeProp * Nmax)
  NmaxControl <- Nmax - NmaxActive

  if (nr && is.null(d)) {
    ## set parameter d for randomly generating look locations
    d <- floor(min(nn - c(0, nn[-nL])) / 2)
  }
  nnr <- nn
  nnrE <- nnE
  nnrF <- nnF
  for (k in 1:ns) { ## simulate a clinical trial ns times
    if (nr && (d > 0)) {
      ## randomly generate look locations
      dd <- sample(-d:d,
        size = nL - 1, replace = TRUE,
        prob = 2^(c(-d:0, rev(-d:(-1))) / 2)
      )
      nnr <- nn + c(dd, 0)
      nnrE <- nnr[nn %in% nnE]
      nnrF <- nnr[nn %in% nnF]
    }

    ## simulate sequence of patients
    isActive <- sample(
      x = rep(c(TRUE, FALSE), c(NmaxActive, NmaxControl)),
      size = Nmax,
      replace = FALSE
    )

    ## simulate sequence of responses
    x <- stats::rbinom(Nmax, 1,
      prob = ifelse(isActive, pE, pS)
    )

    j <- 1
    i <- nnr[j]

    while (is.na(s[k]) && (j <= length(nnr))) {
      ## current data in both arms:
      xActive <- x[which(isActive[1:i])]
      xControl <- x[which(!isActive[1:i])]

      ## compute predictive probability
      if (i %in% nnrF) {
        qL <- 1 - predprobDist(
          x = sum(xActive), n = length(xActive),
          xS = sum(xControl), nS = length(xControl),
          Nmax = NmaxActive, NmaxControl = NmaxControl,
          delta = deltaFu,
          relativeDelta = relativeDelta,
          thetaT = 1 - tFu,
          parE = parE, parS = parS
        )$result

        s[k] <- ifelse(qL >= phiFu, FALSE, NA)
      }
      if (i %in% nnrE) {
        q <- predprobDist(
          x = sum(xActive), n = length(xActive),
          xS = sum(xControl), nS = length(xControl),
          Nmax = NmaxActive, NmaxControl = NmaxControl,
          delta = delta,
          relativeDelta = relativeDelta,
          thetaT = tT,
          parE = parE, parS = parS
        )$result

        ## make the decision
        s[k] <- ifelse(q >= phiU & !(i < Nmax & phiU == 1), ## (1)
          TRUE,
          s[k]
        )
      }

      ## what happens if phiU == 1?
      ## then q >= phiU will always be FALSE, except for the last iteration
      ## when i == Nmax -> then it will be 1 or 0.
      ## If it is 1, then the first condition (1) will be TRUE,
      ## so it will be "success".

      ## sample sizes: total and in both arms
      n[k] <- i
      nActive[k] <- length(xActive)
      nControl[k] <- length(xControl)

      j <- j + 1
      i <- nnr[j]
    }
  }
  oc <- cbind(
    ExpectedN = mean(n),
    ExpectedNactive = mean(nActive),
    ExpectedNcontrol = mean(nControl),
    PrStopEarly = mean(n < Nmax),
    PrEarlyEff = sum(s * (n < Nmax), na.rm = TRUE) / ns,
    PrEarlyFut = sum((1 - s) * (n < Nmax), na.rm = TRUE) / ns,
    PrEfficacy = sum(s, na.rm = TRUE) / ns,
    PrFutility = sum(1 - s, na.rm = TRUE) / ns,
    PrGrayZone = sum(is.na(s) / ns)
  )

  return(list(
    oc = oc, Decision = s, SampleSize = n,
    SampleSizeActive = nActive,
    SampleSizeControl = nControl,
    nn = nn, nnE = nnE, nnF = nnF,
    params = as.list(match.call(expand.dots = FALSE))
  ))
}
```
