---
title: "Design Document for plotOc"
output: html_document
date: "2025-03-05"
editor_options:
  chunk_output_type: console
---
# Purpose 
Plot a bar plot for simulated results of :
- `ocPostprob()` with cases
  1. `length(nnE) = length(nnF)` AND `wiggle = FALSE` ✔️
  2. `length(nnE) = length(nnF)` AND `wiggle = TRUE`. (looks wiggle together)
  3. `length(nnE) =! length(nnF)` AND `wiggle = FALSE`. (looks are different and don't wiggle)
  4. `length(nnE) =! length(nnF)` AND `wiggle = TRUE`. (looks are different and wiggle)
- all other `oc` series such as `ocPredprob()` 

Note these results and definitions :
FALSE = met the futility criteria
TRUE = met the efficacy criteria

which can be improved in future PR, e.g "Go" as factor for "met the efficacy criteria". 

## Important packages

```{r}
library(dplyr)
```

# `ocPostprob()` 

## 1. `ocPostprob()` where `length(nnE) = length(nnF)` AND `wiggle = FALSE`

```{r}
# Here, we only have one Futility but many Efficacy looks or stop.
res1 <- ocPostprob(
  nnE = c(10, 20, 30), truep = 0.40, p0 = 0.20, p1 = 0.30, tL = 0.60, tU = 0.80, parE = c(1, 1),
  sim = 100, wiggle = FALSE, nnF = c(10, 20, 30)
)

res1 # (may need to add wiggled distance)
res1$oc
```

## 2. `ocPostprob()` where `length(nnE) = length(nnF)` AND `wiggle = TRUE`. (looks wiggle together)

```{r}
# Here, we only have one Futility but many Efficacy looks or stop.
res2 <- ocPostprob(
  nnE = c(10, 20, 30), truep = 0.40, p0 = 0.20, p1 = 0.30, tL = 0.60, tU = 0.80, parE = c(1, 1),
  sim = 100, wiggle = TRUE, nnF = c(10, 20, 30)
)

res2 # (may need to add wiggled distance as param)
res2$oc
```

## 3. `ocPostprob()` where `length(nnE) =! length(nnF)` AND `wiggle = FALSE`. (looks are different and don't wiggle)
```{r}
res3 <- ocPostprob(
  nnE = c(10, 20, 30), truep = 0.40, p0 = 0.20, p1 = 0.30, tL = 0.60, tU = 0.80, parE = c(1, 1),
  sim = 100, wiggle = FALSE, nnF = 30
)

res3 # (may need to add wiggled distance)
res3$oc
```

## 4. `ocPostprob()` where `length(nnE) =! length(nnF)` AND `wiggle = TRUE`. (looks are different and wiggle)

```{r}
res4 <- ocPostprob(
  nnE = c(10, 20, 30), truep = 0.40, p0 = 0.20, p1 = 0.30, tL = 0.60, tU = 0.80, parE = c(1, 1),
  sim = 100, wiggle = TRUE, nnF = 30
)

res4 
res4$oc
```

## 5. `ocPredprob()` where decision1 = FALSE
```{r}
res5 <- ocPredprob(
  nnE = c(10, 20),
  truep = 0.6,
  p0 = 0.25,
  p1 = 0.25,
  tT = 0.6,
  tF = 0.6,
  phiU = 0.8,
  phiFu = 0.8,
  parE = c(1, 1),
  sim = 50,
  wiggle = FALSE,
  nnF = c(10, 20),
  decision1 = FALSE
)
res5$oc
```

## 6. `ocPostprobDist()` where decision1 = FALSE
```{r}
res6 <- ocPostprobDist(
  nnE = c(10, 20, 30),
  truep = 0.4,
  deltaE = 0.1,
  deltaF = -0.1,
  tL = 0.6,
  tU = 0.6,
  parE = c(1, 1),
  parS = c(5, 25),
  sim = 50,
  wiggle = TRUE
)
res6$oc
```

## 7. `ocPredprobDist()` where decision1 = FALSE
```{r}
res7 <- ocPredprobDist(
  nnE = c(10, 20, 30),
  truep = 0.40,
  deltaE = 0.10,
  deltaF = 0.10,
  relativeDelta = FALSE,
  tT = 0.6,
  phiU = 0.80,
  phiL = 0.20,
  parE = c(1, 1),
  parS = c(5, 25),
  weights = 1,
  weightsS = 1,
  sim = 50,
  wiggle = FALSE,
  decision1 = TRUE
)
res7$oc
```

### Helper `h_get_dataframe`

```{r}
h_get_dataframe <- function(decision, sample_size, all_looks) {
  df <- data.frame(decision = decision,
                      sample_size = sample_size,
                      look = all_looks)
  # summarise into frequency table
  df <- df %>%  
  group_by(decision, look) %>% 
    summarise(prop = sum(length(decision))/nrow(df)) %>% 
    as_tibble()
  # setting levels of factors
  all_decision <- c(TRUE, FALSE, NA)
  all_looks <- c(10, 20, 30)
  df$decision <- factor(df$decision, levels = all_decision)
  df$look <- factor(df$look, levels = all_looks)
  # data input for plot, such that all variables present, even empty cases
  df <- df %>%
    mutate(decision = forcats::fct_na_value_to_level(decision)) %>%
    complete(decision, look, 
             fill = list(prop = 0)) %>% as.data.frame()
  df

                       }
```

# Example for case 1
```{r}
h_get_dataframe(res1$Decision, sample_size = res1$SampleSize, res1$Looks)
```

# Example for case 2
```{r}
h_get_dataframe(res2$Decision, sample_size = res2$SampleSize, res2$Looks)
```

# Example for case 3
```{r}
h_get_dataframe(res3$Decision, sample_size = res3$SampleSize, res3$Looks)
```

# Example for case 4
```{r}
h_get_dataframe(res4$Decision, sample_size = res4$SampleSize, res4$Looks)
```

# Example for case 5 
```{r}
h_get_dataframe(res5$Decision, sample_size = res5$SampleSize, res5$Looks)
```

# Example for case 6
```{r}
h_get_dataframe(res6$Decision, sample_size = res6$SampleSize, res6$Looks)
```

# Example for case 7
```{r}
h_get_dataframe(res7$Decision, sample_size = res7$SampleSize, res7$Looks)
```


# Testing helper function

```{r}
# h_get_dataframe ----
test_that("h_get_dataframe gives correct data frame when wiggle = TRUE", {
  set.seed(1989)
  res4 <- ocPostprob(
    nnE = c(10, 20, 30), 
    truep = 0.40, 
    p0 = 0.20, 
    p1 = 0.30, 
    tL = 0.60, 
    tU = 0.80, 
    parE = c(1, 1),
    sim = 100, 
    wiggle = TRUE, 
    nnF = 30
  )
  result <- h_get_dataframe(res4$Decision, 
                            sample_size = res4$SampleSize, 
                            res4$Looks)
  expected <- data.frame(decision = 
        structure(c(1L, 1L, 1L, 2L, 2L, 2L, NA, NA, NA), 
                  levels = c("TRUE", "FALSE"), class = "factor"), 
      look = structure(c(1L, 2L, 3L, 1L, 2L, 3L, 1L, 2L, 3L), 
                     levels = c("10","20", "30"), 
                     class = "factor"), 
      prop = c(0.48, 0.18, 0.11, 0, 0, 0.01, 0, 0, 0.22))
  expect_equal(result, expected)
})
```

```{r}
test_that("h_get_dataframe gives correct data frame when wiggle = FALSE", {
  set.seed(1989)
  res3 <- ocPostprob(
    nnE = c(10, 20, 30), 
    truep = 0.40, 
    p0 = 0.20, 
    p1 = 0.30, 
    tL = 0.60, 
    tU = 0.80, 
    parE = c(1, 1),
    sim = 100, 
    wiggle = FALSE, 
    nnF = 30
  )
  result <- h_get_dataframe(res3$Decision, 
                            sample_size = res4$SampleSize, 
                            res4$Looks)
  expected <- data.frame(decision = structure(c(1L, 1L, 1L, 2L, 2L, 2L, NA, NA, NA),
                                              levels = c("TRUE", "FALSE"), 
                                              class = "factor"),
                         look = structure(c(1L, 2L, 3L, 1L, 2L, 3L, 1L, 2L, 3L),
                                          levels = c("10", "20", "30"), 
                                          class = "factor"),
                         prop = c(0.39, 0.13, 0.23, 0, 0, 0,0.09, 0.05, 0.11))
  expect_equal(result, expected)
})
```

## `plotOc`

```{r}
plotOc <- function(decision, sample_size, all_looks, wiggle_status) {
  df <- h_get_dataframe(decision = decision, 
                        sample_size = sample_size, 
                        all_looks = all_looks)
  barplot <-   
    ggplot2::ggplot(df, aes(fill = decision, x = look, y = prop)) +
    ggplot2::geom_bar(position = "dodge", stat = "identity") + 
    ggplot2::ggtitle("Results from simulation : \nProportion of Go/Stop/Grey zone decisions per interim/final analysis") + 
    ggplot2::theme(title = element_text(size = 13)) +
    ggplot2::ylab("percentage") + 
    ggplot2::theme(axis.text.x = element_text(size = 12)) +
    ggplot2::xlab("look (n)") +
    scale_fill_manual(values = c("#009E73", "#FF0046", "lightgrey"),
                      labels = c("Go", "Stop", "Grey zone")) 
    if (wiggle_status) {
        wiggle_warning <- 
          paste("Note that sample sizes slightly differ from the ones labeled")
        barplot + ggplot2::annotate("text", 
                                    label = wiggle_warning, 
                                    x = length(unique(all_looks))/1.5, 
                                    y = 0.65, 
                                    size = 5,
                                    colour = "black") 
      
    } else {
      barplot
    }                               
}
```

### Plot case 1:
```{r}
plotOc(decision = res2$Decision, 
       sample_size = res2$SampleSize,
       all_looks = res2$Looks,
       wiggle_status = res2$params$wiggle)
```
### Plot case 2:
```{r}
plotOc(decision = res1$Decision, 
       sample_size = res1$SampleSize,
       all_looks = res1$Looks,
       wiggle_status = res1$params$wiggle)
```
### Plot case 3:
```{r}
plotOc(decision = res3$Decision, 
       sample_size = res3$SampleSize,
       all_looks = res3$Looks,
       wiggle_status = res3$params$wiggle)
```
### Plot case 4:
```{r}
plotOc(decision = res4$Decision, 
       sample_size = res4$SampleSize,
       all_looks = res4$Looks,
       wiggle_status = res4$params$wiggle)
```

### Plot case 5`ocPredprob()`
```{r}
plotOc(decision = res5$Decision, 
       sample_size = res5$SampleSize,
       all_looks = res5$Looks,
       wiggle_status = res5$params$wiggle)
```

### Plot case 6 `ocPostProbDist``
```{r}
plotOc(decision = res6$Decision, 
       sample_size = res6$SampleSize,
       all_looks = res6$Looks,
       wiggle_status = res6$params$wiggle)
```

### Plot case 7 `ocPredProbDist`
```{r}
plotOc(decision = res7$Decision, 
       sample_size = res7$SampleSize,
       all_looks = res7$Looks,
       wiggle_status = res7$params$wiggle)
```

