---
title: "Design Document for ocPredprob.R"
output: html_document
date: "2024-01-08"
editor_options: 
  chunk_output_type: console
---

## Rule 1
Decision rule 1:The trial is stopped for efficacy if the predictive probability of a
successful trial is larger than phiU, and stopped for futility if it is below
phiL. A trial is successful if after the maximum number of patients the
posterior probability of the treatment having more than p0 response rate is
above tT. Otherwise the decision is "failure". In this case, there is no gray zone.


final decision :

Pr(truep > p0) > tT which represents a successful trial, otherwise if Pr(truep > p0) < tT is not a successful trial.

interim decision:

Predictive probability of Pr( Pr(truep > p0) > tT = Successful trial S) > phiU at interim, then we have an early Go decision

early/interim decision to Stop:

Pr( Pr(truep > p0) > tT = Successful trial S) < phiL

Otherwise we continue the trial, e.g. no early decision made.

## Rule 2
Decision rule 2:A variation can be requested when skipping the argument phiL and utilizing the arguments
p1, tFu & PhiFu. The trial can be stopped for futility if the predictive
probability of an unsuccessful trial is larger than phiFu. In this case, the
decision is "failure" when the posterior probability of having the treatment
response rate at most p1 is above tFu at the maximum number of patients.

at final decision:

Pr (truep < p1) > tFu (if it's too high then it's a failure, here p1 and tFu exist)

at interim:

Pred prob of this failure

Pred prob (failure) > PhiFu (the confidence for failing high, then we stop)

## Design notes :

* the original code does not cater for interim decisions
The code only caters for success critical for Decision 1 and failure for Decision 2
* for main function, we flags based on either decision 1 or 2 (rule 1 or 2)
* 2 helper functions for each decision
* in each, incorporate interim 
* intergrative test with grandma to compare results
* We compare results and maybe we stick to grandma's 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1 Helper function for Decision
    * Decision 1 says:
    * GO = P(truep > p0 | data) => tT, by the way P(truep > p0) > tT = qU
    when the prob of good response is high enough, then we have a successful trial
    * STOP = P(truep > p0 | data ) < tT =  Unsuccessful trial if not successful 
    * EVAL = there is no evaluate BUT there are interims.... in this code
    * interim GO =  P(successful trial at final) > phiU, P(interim_qU > tT) > phiU
    * interim STOP = P(successful trial at final) < phiL, P(interim_qU > tT) < phiL
    * interim EVAL = in between
```{r}
h_get_decision_one_predprob <- function(nnr, truep, p0, parE = c(1, 1), nnE, nnF, tT, phiU, phiL) {
  index_look <- 1
  assert_numeric(nnr)
  size_look <- nnr[index_look]
  all_sizes <- decision <- NA
  response <- stats::rbinom(max(nnr), size = 1, truep)
  assert_numeric(response, lower = 0, upper = 1)
  if (is.na(decision) && index_look <= length(nnr)) { # at interim
    if (size_look %in% nnE) { # GO decision for interim
      interim_qU <- predprob(
        x = sum(x = response[1:size_look]),
        n = size_look,
        Nmax = nnE[length(nnE)],
        p = p0, # p1 only used if decision2 is specified so it does not exist here
        thetaT = tT,
        parE = parE
      )$result
      decision <- ifelse(interim_qU > phiU, TRUE, NA)
    }
    if (size_look %in% nnF) {
      interim_qU <- predprob( # STOP decision for interim
        x = sum(x = response[1:size_look]),
        n = size_look,
        Nmax = nnF[length(nnF)],
        p = p0,
        thetaT = tT,
        parE = parE
      )$result
      decision <- ifelse(interim_qU < phiL, FALSE, decision)
    }
  } else {
    # at final
    if (size_look %in% nnE) { # for efficacy looks
      final_eff_qU <- postprob(
        x = response,
        n = size_look,
        p = p0,
        parE = parE,
        log.p = FALSE
      )
      # it is an final GO if P(response rate for GO > tT ) > phiU
      decision <- ifelse(final_eff_qU > tT, TRUE, FALSE)
    }
    if (size_look %in% nnF) { # for futility looks
      final_fu_qU <- 1 - postprob(
        x = response,
        n = size_look,
        p = p0,
        parE = parE,
        log.p = FALSE
      )$result
      # it is an final STOP if P(response rate for STOP < tT )  < phiL
      decision <- ifelse(final_fu_qU < tT, FALSE, decision)
    }
  }
  all_sizes <- size_look
  index_look <- index_look + 1
  size_look <- nnr[index_look]
  list(
    decision = decision,
    all_sizes = all_sizes
  )
}
```
# Example for decision 1
if nnE = c(10, 20) and nnF =  c(11, 15), and the rules are:

Final:
GO: P (truep > p0 = 0.2) > tT = 0.6
STOP = P(truep > p0 = 0.2) < tT = 0.6 # no grey zone

Interim:
GO: P (successful trial) > phiU = 0.5
STOP: P (successful trial) < phiL = 0.5 # no grey zone

Should be a Go since the truep is assumed to be 0.6 and the threshold for a good response (p0) is 0.2, and the final threshold is only 0.6.

```{r}
h_get_decision_one_predprob(
  nnr = c(10, 20),
  truep = 0.6,
  p0 = 0.2,
  parE = c(1, 1),
  nnE = c(10, 20),
  nnF = c(11, 15),
  tT = 0.6,
  phiU = 0.5,
  phiL = 0.5
)
# (nnr, truep, p0, parE = c(1, 1), nnE, nnF, tT, phiU, phiL)
```

## 2 Helper function for Decision 2
    * Decision 2 says:
    * Final GO = P(truep > p0) < tT =  when the prob of poor response is low, then we have a successful trial 
    * Final STOP = P(truep < p1) > tF:
    when the prob of poor response is high enough, then we have a UNsuccessful trial
    * interim included here .....
    * interim STOP : P (failure at final ) > phiFu
    * interim GO : P ( success at final) > phiU
```{r}
h_get_decision_two_predprob <- function(nnr, truep, p0, p1, parE = c(1, 1), nnE, nnF, tT, tF, phiFu, phiU) {
  index_look <- 1
  assert_numeric(nnr)
  size_look <- nnr[index_look]
  all_sizes <- decision <- NA
  response <- stats::rbinom(max(nnr), size = 1, truep)
  assert_numeric(response, lower = 0, upper = 1)
  if (is.na(decision) && index_look <= length(nnr)) {  # as long s there is no decision...
    if (size_look %in% nnE) {
      interim_qU <- predprob(
        x = sum(x = response[1:size_look]),
        n = size_look,
        Nmax = nnE[length(nnE)],
        p = p0,
        thetaT = tT,
        parE = parE
      )$result
      decision <- ifelse(interim_qU > phiU, TRUE, NA)
    }
    if (size_look %in% nnF) {
      interim_qU <-1- predprob(
        x = sum(x = response[1:size_look]),
        n = size_look,
        Nmax = nnF[length(nnF)],
        p = p1,
        thetaT = tF,
        parE = parE
      )$result
      decision <- ifelse(interim_qU > phiU, FALSE, decision)
    }
  } else {
    if (size_look %in% nnE) { # for efficacy looks
      final_qU <- postprob( # based on all data, the posterior probability of success is a GO when P(truep > p0) > tT
        x = response,
        n = size_look,
        p = truep,
        parE = c(1, 1),
        log.p = FALSE
      )
      decision <- ifelse(final_qU < tT, FALSE, ifelse(interim_qU > phiU, TRUE, decision))
    }
    if (size_look %in% nnF) { # for futility looks
      # based on all data, the posterior probability of success is a STOP when P(truep > p0) < tT
      final_qU <- 1 - postprob(
        x = response,
        n = size_look,
        p = truep,
        parE = c(1, 1),
        log.p = FALSE
      )
      decision <- ifelse(final_qU > tF, FALSE, ifelse(qU > phiU, TRUE, decision))
    }
  }
  all_sizes <- size_look
  index_look <- index_look + 1
  size_look <- nnr[index_look]
  list(
    decision = decision,
    all_sizes = all_sizes
  )
}
```
# Example for decision 2
if nnE = c(10, 20) and nnF =  c(10, 20), and the rules are:

Final:
GO: P (truep > p0 = 0.2) > tT = 0.6
STOP = P(truep < p1 = 0.2) > tF = 0.6 # no grey zone

Interim:
GO: P (successful trial) > phiU = 0.5 
STOP: P (successful trial) > phiFu = 0.5 # no grey zone

when truep = 0.4 and the threshold is lower at 0.2 = p0 = p1.
```{r}
h_get_decision_two_predprob(
  nnr = c(10, 20),
  truep = 0.4,
  p0 = 0.2,
  p1 = 0.2,
  parE = c(1, 1),
  nnE = c(10, 20),
  nnF = c(10, 20),
  tT = 0.6,
  tF = 0.6,
  phiFu = 0.5,
  phiU = 0.5
)

# (nnr, truep, p0, p1, parE = c(1, 1), nnE, nnF, tT, tF, phiFu, phiU)
```

# helper h_get_oc_predprob : the last helper step in the user interfacing function
```{r}
h_get_oc_predprob <- function(all_sizes, nnr, decision, nnrE, nnrF) {
  sim <- length(all_sizes)
  assert_logical(decision, len = sim)
  assert_numeric(all_sizes)
  assert_numeric(nnrE, lower = 0, upper = max(nnrE))
  assert_numeric(nnrF, lower = 0, upper = max(nnrF))
  data.frame(
    ExpectedN = mean(all_sizes, na.rm = TRUE),
    PrStopEarly = mean(all_sizes < max(nnrF), na.rm = TRUE),
    PrEarlyEff = sum(decision * (all_sizes < max(nnrE)), na.rm = TRUE) / sim,
    PrEarlyFut = sum((1 - decision) * (all_sizes < max(nnrF)), na.rm = TRUE) / sim,
    PrEfficacy = sum(decision, na.rm = TRUE) / sim,
    PrFutility = sum(1 - decision, na.rm = TRUE) / sim,
    PrGrayZone = sum(is.na(decision) / sim)
  )
}
```

## new_ocPredprob

```{r cars}
new_ocPredprob <- function(nnE, truep, p0, p1, tT, tF, phiL, phiU, phiFu, parE = c(1, 1),
                           sim = 50000, wiggle = FALSE, nnF = nnE, decision1 = TRUE) {
  nn <- sort(unique(c(nnF, nnE)))
  assert_number(sim, lower = 1, finite = TRUE)
  assert_flag(wiggle)
  if (sim < 50000) {
    warning("Advise to use sim >= 50000 to achieve convergence")
  }
  decision <- vector(length = sim)
  all_sizes <- vector(length = sim)
  for (k in seq_len(sim)) {
    if (length(nn) != 1 && wiggle) {
      dist <- h_get_distance(nn = nn)
      nnr <- h_get_looks(dist = dist, nnE = nnE, nnF = nnF)
      nnrE <- nnr$nnrE
      nnrF <- nnr$nnrF
    } else {
      dist <- 0
      nnrE <- nnE
      nnrF <- nnF
    }
    nnr <- unique(c(nnrE, nnrF))
    if (decision1 == TRUE) { # is.na(decision[k]) &&  in helper 
      tmp <- h_get_decision_one_predprob(
        # (nnr, truep, p0, parE = c(1, 1), nnE, nnF, tT, phiU, phiL)
        nnr = nnr,
        truep = truep,
        p0 = p0,
        parE = parE,
        nnE = nnrE,
        nnF = nnrF,
        tT = tT,
        phiU = phiU,
        phiL = phiL
      )
      decision[k] <- tmp$decision
      all_sizes[k] <- tmp$all_sizes
    } else {
      tmp <- h_get_decision_two_predprob(
        # (nnr, truep, p0, p1, parE = c(1, 1), nnE, nnF, tT, tF, phiFu, phiU)
        nnr = nnr,
        truep = truep,
        p0 = p0,
        p1 = p1,
        parE = parE,
        nnE = nnrE,
        nnF = nnrF,
        tT = tT,
        tF = tF,
        phiFu = phiFu,
        phiU = phiU
      )
      decision[k] <- tmp$decision
      all_sizes[k] <- tmp$all_sizes
    }
  }
  oc <- h_get_oc_predprob(all_sizes = all_sizes, nnr = nnr, decision = decision, nnrE = nnrE, nnrF = nnrF)
  list(
    oc = oc,
    Decision = decision,
    SampleSize = all_sizes,
    union_nn = nnr,
    input_nnE = nnE,
    input_nnF = nnF,
    wiggled_nnE = nnrE,
    wiggled_nnF = nnrF,
    wiggle_dist = dist,
    params = as.list(match.call(expand.dots = FALSE))
  )
}
```
# Example for dec 1
```{r, echo=FALSE}
dec1 <- new_ocPredprob(
  nnE = c(10, 20, 30),
  truep = 0.4,
  p0 = 0.25,
  p1 = NA,
  tT = 0.6,
  tF = 1 - 0.6,
  phiL = 0.2,
  phiU = 0.10,
  phiFu = NA,
  parE = c(1, 1),
  sim = 100,
  wiggle = FALSE,
  nnF = c(10, 22, 25),
  decision1 = TRUE
)
dec1
# (nnE, truep, p0, p1, tT, tFu, phiL, phiU, phiFu, parE = c(1, 1),
# sim = 50000, wiggle = FALSE, nnF = nnE, decision1 = TRUE)
```
# Example for dec 2
```{r, echo=FALSE}
dec2 <- new_ocPredprob(
  nnE = c(10, 20, 30),
  truep = 0.4,
  p0 = 0.25,
  p1 = 0.25,
  tT = 0.6,
  tF = 1 - 0.6,
  phiL = 0.2,
  phiU = 0.10,
  phiFu = 0.1,
  sim = 100,
  wiggle = FALSE,
  nnF = c(10, 22, 25),
  decision1 = FALSE
)
dec2
# (nnE, truep, p0, p1, tT, tFu, phiL, phiU, phiFu, parE = c(1, 1),
# sim = 50000, wiggle = FALSE, nnF = nnE, decision1 = TRUE)
```

## Integration test: grandma vs new
Example: To examine operating characteristics of a phase Ib design with 2 interim analyses
(sample size @ 10, 20 using predictive probability) and one final analysis (sample size @ 30);

* design details
*  multiple looks @ 10, 20, 30 patients
*  True response rate of the treatment group = 40%
*  control response rate = 25%

*  denote treatment response rate as RRS
*  trial success is defined as: P(RRS > control) > tT

 * Decision 1 says:
    * GO = P(truep > p0 | data ) > tT, by the way P(truep > p0) > tT = qU
    when the prob of good response is high enough, then we have a successful trial
    * STOP = P(truep > p0  | data) < tT =  Unsuccessful trial if not successful 
    * EVAL = there is no evaluate BUT there are interims.... in this code
    * interim GO =  P(successful trial | interim data ) > phiU, P(qU > tT) > phiU
    * interim STOP = P(successful trial | interim data) > phiL, P(qU > tT) < phiL
    * interim EVAL = in between

therefore in this example:

 GO = P(truep => p0 = 0.4) > tT = 0.6
 STOP =  P(truep > p0 = 0.4) < tT = 0.6
 interim GO =  P(successful trial) > phiU = 0.8, P(interim_qU > tT) > phiU
 interim STOP = P(successful trial) > phiL = 0.3, P(interim_qU > tT) < phiL
 interim EVAL = in between
 

```{r}
set.seed(20)
result <- ocPredprob(
  nn = c(10, 20, 30),
  p = 0.4, # truep
  p0 = 0.25,
  tT = 0.6,
  phiL = 0.3,
  phiU = 0.8,
  parE = c(1, 1),
  ns = 50
)$oc
result 
expected <- new_ocPredprob(
  nnE = c(10, 20, 30),
  truep = 0.4,
  p0 = 0.25,
  p1 = NA,
  tT = 0.6,
  phiL = 0.3,
  phiU = 0.8,
  parE = c(1, 1),
  sim = 50,
  wiggle = FALSE,
  nnF = c(10, 20, 30),
  decision1 = TRUE
)$oc
expected
# >   result
#      ExpectedN PrStopEarly PrEarlyEff PrEarlyFut PrEfficacy PrFutility PrGrayZone
# [1,]    13.875     0.94906    0.87224    0.07682    0.91414    0.08586          0
# > expected
# $oc
#   ExpectedN PrStopEarly PrEarlyEff PrEarlyFut PrEfficacy PrFutility
# 1        10           1    0.61876    0.04592    0.61876    0.04592


```


```{r}
set.seed(1989)
result <- ocPredprob(
  nn = c(10, 20, 30),
  p = 0.4, # truep
  p0 = 0.25,
  p1 = 0.30,
  tT = 0.6,
  phiL = 0.3,
  phiU = 0.8,
  parE = c(1, 1),
  ns = 50
)$oc
result
# > result
#      ExpectedN PrStopEarly PrEarlyEff PrEarlyFut PrEfficacy PrFutility PrGrayZone
# [1,]    12.141           1     0.7942     0.2058     0.7942     0.2058          0
expected <- new_ocPredprob(
  nnE = c(10, 20, 30),
  truep = 0.4,
  p0 = 0.25,
  p1 = 0.30,
  tT = 0.6,
  tF = 0.5,
  phiL = 0.3,
  phiU = 0.8,
  parE = c(1, 1),
  sim = 50,
  wiggle = FALSE,
  nnF = c(10, 20, 30),
  decision1 = FALSE
)$oc
expected
# > expected
#   ExpectedN PrStopEarly PrEarlyEff PrEarlyFut PrEfficacy PrFutility
# 1        10           1    0.61456    0.38544    0.61456    0.38544

```

# Are we getting the same results ? We simulate sim number of trials to check if in decision 1 and decision 2, we are getting the same results
1. Run ten simulations of decision 1, compare it with the grandma function
2. Run ten simulations of decision 2, compare it with the grandma function
```{r}
sn_ocPredprob <- function(nnE, nnF, truep, p0, p1, tT, tF, phiU, phiL, phiFu, sim, wiggle = FALSE, decision1 = TRUE, parE = c(1,1)) {
   nn <- sort(unique(c(nnF, nnE)))
   decision <- vector(length = sim)
   all_sizes <- vector(length = sim)
   for (k in seq_len(sim)) {
    if (length(nn) != 1 && wiggle) {
      dist <- h_get_distance(nn = nn)
      nnr <- h_get_looks(dist = dist, nnE = nnE, nnF = nnF)
      nnrE <- nnr$nnrE
      nnrF <- nnr$nnrF
    } else {
      dist <- 0
      nnrE <- nnE
      nnrF <- nnF
    }
    nnr <- unique(c(nnrE, nnrF))
    if (decision1 == TRUE) { 
      tmp <- h_get_decision_one_predprob(
        nnr = nnr,
        truep = truep,
        p0 = p0,
        parE = parE,
        nnE = nnrE,
        nnF = nnrF,
        tT = tT,
        phiU = phiU,
        phiL = phiL
      )
      decision[k] <- tmp$decision
      all_sizes[k] <- tmp$all_sizes
    } else {
      tmp <- h_get_decision_two_predprob(
        nnr = nnr,
        truep = truep,
        p0 = p0,
        p1 = p1,
        parE = parE,
        nnE = nnrE,
        nnF = nnrF,
        tT = tT,
        tF = tF,
        phiFu = phiFu,
        phiU = phiU
      )
      decision[k] <- tmp$decision
      all_sizes[k] <- tmp$all_sizes
    }
   }
   oc <- h_get_oc_predprob(all_sizes = all_sizes, nnr = nnr, decision = decision, nnrE = nnrE, nnrF = nnrF)
   list(decision = decision, all_sizes = all_sizes, oc = oc)
}

g_ocPredprob <- function(nn, nnE, nnF, ns, d = NULL, p, p0, p1, tT, tFu, parE, phiFu, phiU, nr = FALSE) {
  nnE <- sort(nn)
  nnF <- sort(nnF)
  s <- rep(NA, ns)
  n <- s
  nn <- sort(unique(c(nnF, nnE)))
  nL <- length(nn)
  Nstart <- nn[1]
  Nmax <- nn[nL]
  if (nr && is.null(d)) {
    # set parameter d for randomly generating look locations
    d <- floor(min(nn - c(0, nn[-nL])) / 2)
  }
  nnr <- nn
  nnrE <- nnE
  nnrF <- nnF
  for (k in 1:ns) {
    # simulate a clinical trial ns times
    if (nr && (d > 0)) {
      # randomly generate look locations
      dd <- sample(-d:d,
                   size = nL - 1, replace = TRUE,
                   prob = 2^(c(-d:0, rev(-d:(-1))) / 2)
      )
      nnr <- nn + c(dd, 0)
      
      
      nnrE <- nnr[nn %in% nnE]
      nnrF <- nnr[nn %in% nnF]
    }
    x <- stats::rbinom(Nmax, 1, p)
    j <- 1
    i <- nnr[j]
    for (k in 1:ns) {
      # simulate a clinical trial ns times
      if (nr && (d > 0)) {
        # randomly generate look locations
        dd <- sample(-d:d,
                     size = nL - 1, replace = TRUE,
                     prob = 2^(c(-d:0, rev(-d:(-1))) / 2)
        )
        nnr <- nn + c(dd, 0)
        nnrE <- nnr[nn %in% nnE]
        nnrF <- nnr[nn %in% nnF]
      }
      x <- stats::rbinom(Nmax, 1, p)
      j <- 1
      i <- nnr[j]
      while (is.na(s[k]) && (j <= length(nnr))) {
        if (i %in% nnrF) {
          qL <- 1 - predprob(x = sum(x[1:i]), n = i, Nmax = Nmax, p = p1, thetaT = 1 - tFu, parE = parE)$result
          s[k] <- ifelse(qL > phiFu, FALSE, NA)
        }
        if (i %in% nnrE) {
          q <- predprob(x = sum(x[1:i]), n = i, Nmax = Nmax, p = p0, thetaT = tT, parE = parE)$result
          s[k] <- ifelse(q >= phiU & !(i < Nmax & phiU == 1), TRUE, s[k])
        }
        n[k] <- i
        j <- j + 1
        i <- nnr[j]
      }
       oc <- h_get_oc_predprob(all_sizes = n, nnr = nnr, decision = d, nnrE = nnrE, nnrF = nnrF)
    }
  }
}
```
# Decision 1 simulations vs grandma's decision 1
```{r}
set.seed(2024)
tt <- list(nnE = c(10, 20, 30) , nnF = c(10, 20, 30), truep = 0.2, p0 = 0.3, p1 = 0.2, tT = 0.5, tF = 0.5, phiU = 0.6, phiL = 0.7, phiFu = 0.5, sim = 100, decision1 = TRUE, parE = c(1,1))
sn_ocPredprob(nnE = tt$nnE, nnF = tt$nnF, truep = tt$truep, p0 = tt$p0, p1 = tt$p1, tT = tt$tT, tF = tt$tF, phiU = tt$phiU, phiL = tt$phiL, phiFu = tt$phiFu, sim = tt$sim, decision1 = tt$decision1, parE = tt$parE)

g_ocPredprob(nn = tt$nnE, nnE= tt$nnE, nnF = tt$nnF, ns = tt$sim, d = NULL, p = tt$truep, p0 = tt$p0, p1 = tt$p1, tT = tt$tT, tFu = tt$tF, parE = tt$parE, phiFu = tt$tF, phiU = tt$phiU, nr = FALSE)
#(nn, nnE, nnF, ns, d = NULL, p, p0, p1, tT, tFu, parE, phiFu, phiU, nr = FALSE)
```

# Decision 2 simulations vs
