---
title: "Design document for ocPostprobDist.R"
output: html_document
date: "2023-12-19"
---
# Input parameters from example file :

design details
multiple looks @ 10, 20, 30 patients
True response rate of the treatment group=0.4

stop for efficacy (deltaE): P(pE > pS + deltaE) >tU
stop for futility (deltaF): P(pE < pS - deltaF) >tL
where pE is the response rate of treatment group, pS is the response rate of
control group. Both of them are random variables.

Design notes :

Three helper files assist. Two from ocPostprob.R. The only one specialised for ocPostprobDist is `h_get_decisionDist` which gives the end result.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

input <- list(
  nn = c(10, 20, 30),
  nnE = c(10, 20, 30),
  nnF = c(10, 20, 30),
  p0 = 0.2,
  p1 = 0.3,
  parE = c(1, 1),
  parS = c(5, 25),
  truep = 0.4,
  tU = 0.8,
  tL = 0.6,
  wiggle = TRUE,
  sim = 100,
  randomdist = NULL,
  relativeDelta = FALSE,
  deltaE = 0.1,
  deltaF = -0.1
)
```


# h_get_decision as helper function

We use the `h_get_decision` from `postprob.R` to evaluate the posteriors and create an "array" of "vector" of decision results. `h_get_decision` was adapted by incorporating `postprobDist()` instead of `postprob()` to evaluate this posterior. The methodology is from Thall & Simon 1994 (Biometrics).

```{r}
h_get_decisionDist <- function(nnr,
                               nnrE,
                               nnrF,
                               truep,
                               parE = c(1, 1),
                               parS = c(1, 1),
                               tL,
                               tU,
                               deltaF,
                               deltaE,
                               relativeDelta) {
  index_look <- 1
  size_look <- nnr[index_look]
  all_sizes <- decision <- NA
  response <- stats::rbinom(max(nnr), size = 1, truep)

  while (is.na(decision) && index_look <= length(nnr)) {
    if (size_look %in% nnrF) {
      qL <- postprobDist(
        x = 0,
        n = 0,
        xS = sum(response[1:size_look]),
        nS = size_look,
        delta = deltaF,
        relativeDelta = relativeDelta,
        parE = parS,
        parS = parE
      )
      decision <- ifelse(qL >= tL, FALSE, NA)
    }
    if (size_look %in% nnrE) {
      qU <- postprobDist(
        x = sum(response[1:size_look]),
        n = size_look,
        xS = 0,
        nS = 0,
        delta = deltaE,
        relativeDelta = relativeDelta,
        parE = parE,
        parS = parS
      )
      decision <- ifelse(qU < tU, decision, TRUE)
    }
    all_sizes <- size_look
    index_look <- index_look + 1
    size_look <- nnr[index_look]
  }
  list(
    decision = decision,
    all_sizes = all_sizes
  )
}
```

# Example to test `h_get_decisionDist()`.
```{r}
example_dist <- h_get_distance(nn = c(10, 20, 30))
looks <- h_get_looks(dist = example_dist, nnE = c(10, 20, 30), nnF = c(10, 20, 30))
looks_nnrE <- looks$nnrE
looks_nnrF <- looks$nnrF

# input values from above arguments
h_get_decisionDist(
  nnr = c(10, 20, 30),
  nnrE = looks_nnrE,
  nnrF = looks_nnrF,
  truep = 0.4,
  parE = c(1, 1),
  tL = 0.8,
  tU = 0.6,
  deltaF = 0.1,
  deltaE = 0.2,
  relativeDelta = TRUE
)
```

## Main Function
```{r cars}
ocPostprobDist <- function(nn,
                           truep,
                           deltaE,
                           deltaF,
                           relativeDelta = FALSE,
                           tL,
                           tU,
                           parE = c(a = 1, b = 1),
                           parS = c(a = 1, b = 1),
                           wiggle = TRUE,
                           sim = 50000,
                           nnF = nn,
                           randomdist = TRUE) {
  assert_numeric(nn, min.len = 1, lower = 1, upper = max(nn), any.missing = FALSE)
  assert_number(truep, lower = 0, upper = 1)
  assert_number(deltaE, upper = 1, finite = TRUE)
  assert_number(deltaF, upper = 1, finite = TRUE)
  assert_flag(relativeDelta)
  assert_number(tL, lower = 0, upper = 1)
  assert_number(tU, lower = 0, upper = 1)
  assert_numeric(parE, lower = 0, finite = TRUE, any.missing = FALSE)
  assert_numeric(parS, lower = 0, finite = TRUE, any.missing = FALSE)
  assert_number(sim, lower = 100, finite = TRUE)
  assert_flag(wiggle)
  assert_numeric(nnF, min.len = 1, any.missing = FALSE)
  assert_flag(randomdist)

  if (sim < 50000) {
    warning("Advise to use sim >= 50000 to achieve convergence")
  }
  decision <- vector(length = sim)
  all_sizes <- vector(length = sim)

  nnE <- sort(nn)
  nnF <- sort(nnF)
  nn <- sort(unique(c(nnF, nnE)))
  nL <- length(nn)
  Nstart <- nn[1]
  Nmax <- nn[nL]

  # simulate a clinical trial sim times
  for (k in seq_len(sim)) {
    if (length(nn) != 1 && wiggle && randomdist) {
      # randomly generate look locations
      dist <- h_get_distance(nn = nn)
      nnr <- h_get_looks(dist = dist, nnE = nnE, nnF = nnF) # we generate sim number of looks
      nnrE <- nnr$nnrE
      nnrF <- nnr$nnrF
    } else {
      nnrE <- nnE
      nnrF <- nnF
    }
    nnr <- unique(c(nnrE, nnrF))
    tmp <- h_get_decisionDist(
      nnr = nnr,
      nnrE = nnrE,
      nnrF = nnrF,
      truep = truep,
      parE = c(1, 1),
      tL = tL,
      tU = tU,
      deltaE = deltaE,
      deltaF = deltaF,
      relativeDelta = relativeDelta
    )
    decision[k] <- tmp$decision
    all_sizes[k] <- tmp$all_sizes
  }
  oc <- h_get_oc(all_sizes = all_sizes, nnr = nnr, decision = decision, nnrE = nnrE, nnrF = nnrF)
  list(
    oc = oc,
    Decision = decision,
    SampleSize = all_sizes,
    union_nn = nnr,
    input_nnE = nnE,
    input_nnF = nnF,
    wiggled_nnE = nnrE,
    wiggled_nnF = nnrF,
    wiggle_dist = dist,
    params = as.list(match.call(expand.dots = FALSE))
  )
}
```

## Example
```{r pressure, echo=FALSE}
res3 <- ocPostprobDist(
  nn = c(10, 20, 30),
  truep = 0.4,
  deltaE = 0.1,
  deltaF = 0.1,
  relativeDelta = FALSE,
  tL = 0.6,
  tU = 0.6,
  parE = c(a = 1, b = 1),
  parS = c(a = 1, b = 1),
  sim = 100,
  randomdist = TRUE,
  wiggle = TRUE
)

res3$oc
res3
```

```{r}
res2 <- ocPostprobDist(
  nn = c(10, 20, 30),
  truep = 0.4,
  deltaE = 0.1,
  deltaF = -0.1,
  relativeDelta = FALSE,
  tL = 0.6,
  tU = 0.6,
  parE = c(1, 1),
  parS = c(5, 25),
  sim = 100,
  randomdist = TRUE,
  wiggle = TRUE
)

res2
```
