---
title: "Design Doc for plotBounds"
format: revealjs
editor: visual
---

Statisticians can display boundaries of Go, Stop and GreyZone decisions graphically in terms of response rates required by desired samples sizes or looks. This is such purpose of `plotBounds`. Note that for the predictive probability case, decision 1 rules are chosen. See `ocPredprob` for rules. 

# packages that aid

```{r}
library(flextable)
library(ggplot2)
```

# input

```{r}
boundsPostprob(
    looks = c(10, 20, 30, 40),
    p0 = 0.20,
    tL = 0.10,
    tU = 0.90,
    parE = c(1, 1)
  )
```

```{r}
boundsPredprob(
    looks = c(10, 20, 30, 40),
    p0 = 0.20,
    tT = 0.80,
    phiL = 0.10,
    phiU = 0.90,
  )
```

# vintage function

```{r}
vintage_plotBounds <- function(
  z,
  area = TRUE,
  grid = TRUE,
  yt = "x",
  add = FALSE,
  cols = c("#009E73", "#FF0046", "#7F55B1", "#F0A04B"),
  lwds = c(3, 3),
  ltype = "l",
  lpch = 16,
  lcex = 1,
  gy = 20
) {
  n <- nrow(z)
  nmin <- min(z$looks)
  nmax <- max(z$looks)
  if (yt == "x") {
    z1 <- z$xL
    z2 <- z$xU
    yU <- nmax
    yU2 <- nmin
    ylabel <- "Number of Responses"
    gridy <- seq(0, yU, by = floor(yU / gy))
  } else if (yt == "p") {
    z1 <- z$pL
    z2 <- z$pU
    yU <- 1
    yU2 <- 1
    ylabel <- "Response Rate"
    gridy <- seq(0, yU, by = yU / gy)
  } else {
    stop("yt can only be x or p")
  }
  if (add) {
    graphics::lines(
      z$looks,
      z2,
      lwd = lwds[1],
      col = cols[3],
      type = ltype,
      pch = lpch,
      cex = lcex
    )
    graphics::lines(
      z$looks,
      z1,
      lwd = lwds[2],
      col = cols[4],
      type = ltype,
      pch = lpch,
      cex = lcex
    )
    return(invisible())
  }
  graphics::plot(
    z$looks,
    rep(0, n),
    xlim = c(0, max(z$looks)),
    ylim = c(0, yU),
    type = "n",
    xlab = "n",
    ylab = ylabel
  )
  if (grid) {
    graphics::abline(h = gridy, col = "gray")
  }
  if (area) {
    graphics::polygon(
      c(z$looks, nmax, nmin),
      c(z2, yU, yU2),
      lwd = lwds[1],
      col = cols[1],
      border = cols[1]
    )
    graphics::polygon(
      c(z$looks, nmax, nmin),
      c(z1, 0, 0),
      lwd = lwds[2],
      col = cols[2],
      border = cols[2]
    )
  } else {
    graphics::lines(
      z$looks,
      z2,
      lwd = lwds[1],
      col = cols[1],
      type = ltype,
      pch = lpch,
      cex = lcex
    )
    graphics::lines(
      z$looks,
      z1,
      lwd = lwds[2],
      col = cols[2],
      type = ltype,
      pch = lpch,
      cex = lcex
    )
  }
  return(invisible())
}

```

# vintage example that should work with new function

```{r}
plotBounds(
  boundsPostprob(
    looks = c(10, 20, 30, 40),
    p0 = 0.20,
    tL = 0.10,
    tU = 0.90,
    parE = c(1, 1)
  ),
  yt = "p",
  add = TRUE
)

```

# vintage function

```{r}
vintage_plotBounds(
  boundsPredprob(
    looks = c(10, 20, 30, 40),
    p0 = 0.20,
    tT = 0.80,
    phiL = 0.10,
    phiU = 0.90,
  ),
  yt = "x"
)
```


```{r}
new_plotBounds <- function(coords, 
                           postprob = TRUE, 
                           go_colour = "#1B9E77",
                           stop_colour = "#E41A1C",
                           alpha = 1) {
  go_data <- data.frame(
    n = coords$looks,
    ymin_go = coords$xU,  
    ymax_go = coords$looks 
  )
  stop_data <- data.frame(
    n = coords$looks,
    ymin_stop = c(0, 0, 0, 0),     
    ymax_stop = coords$xL      
  )
  plot <- ggplot2::ggplot() +
    ggplot2::geom_ribbon(
      data = stop_data,
      ggplot2::aes(x = n, ymin = ymin_stop, ymax = ymax_stop),
      fill = stop_colour, alpha = alpha
    ) +
    ggplot2::geom_ribbon(
      data = go_data,
      ggplot2::aes(x = n, ymin = ymin_go, ymax = ymax_go),
      fill = go_colour , alpha = alpha
    ) +
    ggplot2::scale_x_continuous(
      name = "n",
      breaks = go_data$n
    ) +
    ggplot2::scale_y_continuous(
      breaks = seq(from = 0, to = max(go_data$n), by = 2)
    ) +
     ggplot2::geom_point(
    data = go_data,
     ggplot2::aes(x = n, y = ymin_go),
    size = 3,       
    shape = 16,
    colour = "#16C47F", 
  ) +
     ggplot2::geom_point(
    data = stop_data,
     ggplot2::aes(x = n, y = ymax_stop),
    size = 3,       
    shape = 16,
    colour = "#FF748B", #"#DC3C22",FF748B 
  ) +
    ggplot2::xlab("Sample Size (n)") +
    ggplot2::ylab("Number of Responders") +
    ggplot2::theme_minimal() 
  
    plot + ggplot2::ggtitle(
      "Boundaries of Go and Stop Responders per sample size")
}
```

# creating polygons with new function with `boundsPredprob()`

The following output show that important coordinates needed as input to create these polygons. 

As a premise, we establish the criteria:

Interim
Go : P(success) > 90 %
Stop : P(success) > 10 %

Final
Go : P(RR > 0.20 | data) > 80 %
Stop : P(RR > 0.20 | data) < 10 %

The criteria for interim analyses are from predictive probability calculation.

The criteria for final analyses are from posterior probability calculation. 

which translates to boundaries :
When n = 10, 4 responders and above is a Go, 0 responders and below is a stop.
When n = 20, 7 responders and above is a Go, 2 responders and below is a stop
When n = 30, 9 responders and above is a Go, 5 responders and below is a stop
When n = 40, 10 responders and above is a Go, 9 responders and below is a stop

```{r}
coords <- boundsPredprob(
    looks = c(10, 20, 30, 40),
    p0 = 0.20,
    tT = 0.80,
    phiL = 0.10,
    phiU = 0.90,
  )
print(coords)
```


```{r}
coords <- boundsPredprob(
    looks = c(10, 20, 30, 40),
    p0 = 0.20,
    tT = 0.80,
    phiL = 0.10,
    phiU = 0.90,
  )
new_plotBounds(coords)
```

# creating polygons with new function with `boundsPostprob()`

The following output show that important coordinates needed as input to create these polygons. 

As a premise, we establish the criteria:
Go : P(RR > 0.20 | data) > 90 %
Stop : P(RR > 0.20 | data) > 10 %
which translates to boundaries :
When n = 10, 4 responders and above is a Go, 3 responders and below is a stop.
When n = 20, 7 responders and above is a Go, 6 responders and below is a stop
When n = 30, 9 responders and above is a Go, 8 responders and below is a stop
When n = 40, 12 responders and above is a Go, 11 responders and below is a stop


```{r}
coords1 <- boundsPostprob(
    looks = c(10, 20, 30, 40),
    p0 = 0.20,
    tL = 0.10,
    tU = 0.90,
    parE = c(1, 1)
  )
new_plotBounds(coords1)
```
















