---
title: "Design_doc_ocPredprobDist"
output: html_document
date: "2024-02-15"
---
For formulas and expressions for predictive probability, refer to [predprobDist()].

The rules for Stop, Go and Grey Zone (where applicable) are the same as in [ocPredprob()].
However when compared to [ocPredprob()], where the predictive probability of `response rate > p0` was used to reach decisions, here
the `response rate - p0 > 0` is replaced by `response rate - p0 > delta`, and the reason is that `p0` is now also uncertain, which is quantified by a posterior distribution.

To therefore understand the margin of difference between the experimental and control group, refer
to the two ways of calculating delta in [postprobDist()]. The margin `delta` from [postprobDist()] is no longer used however where a Go decision is evaluated, the margin `deltaE` is instead employed, and where a Stop decision is evaluated, the margin `deltaF` is employed.

As mentioned, the rules of [ocPredprobDist()] and [ocPredprob()] for both Decision 1 and Decision 2 are the same, and the margin of difference differ depending on whether it is a Go or Stop decision. We highlight here another distinction between [ocPredprob()] and [ocPredprobDist()], seen in the evaluation for the final futility look in [ocPredprobDist()] :

- `Pr(P_E < p1 + deltaF | data) > tF` for the absolute case
- `Pr(P_E < (1-p1)*deltaF | data) > tF` for the relative case

# input
```{r}
input <- list(
  nnr = c(10, 20, 30),
  nnE = c(10, 20, 30),
  nnF = c(10, 20, 30),
  x = 16,
  n = 23,
  xS = 5,
  nS = 10,
  Nmax = 40,
  NmaxControl = 20,
  delta = 0,
  thetaT = 0.6,
  parE = t(c(1, 1)),
  parS = t(c(1, 1)),
  weights = 1,
  weightsS = 1,
  relativeDelta = FALSE,
  phiU = 0.8,
  phiL = 0.6,
  phiFu = 0.6,
  truep = 0.4,
  tT = 0.6,
  deltaE = 0.1,
  deltaF = 0.1,
  p0 = 0.5,
  p1 = 0.5
)
```

## h_get_decision_one...helper function for decision 1

```{r}
h_get_decision_one_predprobDist <- function(nnE,
                                            nnF,
                                            nnr,
                                            truep,
                                            xS,
                                            nS,
                                            parE = c(1, 1),
                                            parS = c(1, 1),
                                            tT,
                                            phiU,
                                            phiL,
                                            deltaE,
                                            deltaF,
                                            weights,
                                            weightsS,
                                            relativeDelta = FALSE) {
  assert_numeric(nnE, lower = 1, sorted = TRUE, any.missing = FALSE)
  assert_numeric(nnF, lower = 1, sorted = TRUE, any.missing = FALSE)
  assert_numeric(nnr, lower = 1, sorted = TRUE)
  assert_number(truep, lower = 0, upper = 1)
  assert_number(xS, finite = TRUE)
  assert_number(nS, finite = TRUE)
  assert_numeric(parE, min.len = 2, any.missing = FALSE)
  assert_numeric(parS, min.len = 2, any.missing = FALSE)
  assert_number(tT, lower = 0, upper = 1)
  assert_number(phiU, lower = 0, upper = 1)
  assert_number(phiL, lower = 0, upper = 1)
  assert_number(deltaE, lower = 0, upper = 1)
  assert_number(deltaF, lower = 0, upper = 1)
  assert_number(weights, lower = 1, finite = TRUE)
  assert_number(weightsS, lower = 1, finite = TRUE)
  assert_flag(relativeDelta)

  index_look <- 1
  Nmax <- max(nnr)
  NmaxControl <- max(nnF)
  decision <- NA
  response <- stats::rbinom(Nmax, size = 1, prob = truep)

  ### Decision 1:
  # The criteria for Decision 1 for Interim looks are :
  # interim GO =  P(successful trial at final) > phiU
  # interim STOP = P(successful trial at final) < phiL
  while (is.na(decision) && index_look < length(nnr)) {
    size_look <- nnr[index_look]
    if (size_look %in% nnE) {
      interim_qU <- predprobDist(
        x = sum(response[1:size_look]),
        n = size_look,
        xS = xS,
        nS = nS,
        Nmax = Nmax,
        NmaxControl = NmaxControl,
        delta = deltaE,
        relativeDelta = relativeDelta,
        parE = parE,
        weights = weights,
        parS = parS,
        weightsS = weightsS,
        thetaT = tT
      )$result
      decision <- ifelse(interim_qU > phiU, TRUE, decision)
    }
    if (size_look %in% nnF) {
      interim_qU <- predprobDist(
        x = sum(response[1:size_look]),
        n = size_look,
        xS = xS,
        nS = nS,
        Nmax = Nmax,
        NmaxControl = NmaxControl,
        delta = deltaF,
        relativeDelta = relativeDelta,
        parE = parE,
        weights = weights,
        parS = parS,
        weightsS = weightsS,
        thetaT = tT
      )$result
      decision <- ifelse(interim_qU < phiL, FALSE, decision)
    }
    index_look <- index_look + 1
  }
  # The criteria for Decision 1 for Final looks are:
  #- Final GO = P( RR > p0 + delta | data) => tT
  #- Final STOP = P(RR > p0 - delta | data ) < tT
  if (is.na(decision)) {
    size_look <- nnr[index_look]
    if (size_look %in% nnE) {
      final_eff_qU <- postprobDist(
        x = sum(response[1:size_look]),
        n = size_look,
        xS = xS,
        nS = nS,
        delta = deltaE,
        relativeDelta = relativeDelta,
        parE = parE,
        weights = weights, # for activeBetamixPost
        parS = parS,
        weightsS = weightsS # for controlBetamixPost
      )
    }
    decision <- ifelse(final_eff_qU >= tT, TRUE, NA)
  }
  if (size_look %in% nnF) {
    final_fu_qU <- postprobDist(
      x = sum(response[1:size_look]),
      n = size_look,
      xS = xS,
      nS = nS,
      delta = deltaF,
      relativeDelta = relativeDelta,
      parE = parE,
      weights = weights, # for activeBetamixPost
      parS = parS,
      weightsS = weightsS # for controlBetamixPost
    )
    decision <- ifelse(final_fu_qU < tT, FALSE, decision)
  }
  list(
    decision = decision,
    all_sizes = size_look
  )
}
```

# Example for h_get_decision_one_predprobDist
```{r}
example_dist <- h_get_distance(nn = c(10, 20, 30))
looks <- h_get_looks(dist = example_dist, nnE = c(10, 20, 30), nnF = c(10, 20, 30))
looks_nnrE <- looks$nnrE
looks_nnrF <- looks$nnrF

# input values from above arguments
h_get_decision_one_predprobDist(
  nnE = looks$nnrE,
  nnF = looks$nnrF,
  nnr = input$nnr,
  truep = input$truep,
  xS = input$xS,
  nS = input$nS,
  parE = c(1, 1),
  parS = c(1, 1),
  tT = input$tT,
  phiU = input$phiU,
  phiL = input$phiL,
  deltaE = input$deltaE,
  deltaF = input$deltaF,
  weights = c(input$weights),
  weightsS = c(input$weightsS),
  relativeDelta = FALSE
)
```
## h_get_decision_two_predprobDist helper function for decision 2

```{r}
h_get_decision_two_predprobDist <- function(
    nnE,
    nnF,
    nnr,
    truep,
    xS,
    nS,
    parE = c(1, 1),
    parS = c(1, 1),
    tT,
    tF,
    phiU,
    phiL,
    phiFu,
    deltaE,
    deltaF,
    weights,
    weightsS,
    relativeDelta) {
  assert_numeric(nnE, lower = 1, sorted = TRUE, any.missing = FALSE, )
  assert_numeric(nnF, lower = 1, sorted = TRUE, any.missing = FALSE, )
  assert_numeric(nnr, lower = 1, sorted = TRUE, any.missing = FALSE, )
  assert_number(truep, lower = 0, upper = 1)
  assert_numeric(xS, sorted = TRUE)
  assert_numeric(nS, sorted = TRUE)
  assert_numeric(parE, min.len = 2, any.missing = FALSE)
  assert_numeric(parS, min.len = 2, any.missing = FALSE)
  assert_number(tT, lower = 0, upper = 1)
  assert_number(phiU, lower = 0, upper = 1)
  assert_number(phiL, lower = 0, upper = 1)
  assert_number(phiFu, lower = 0, upper = 1)
  assert_number(deltaE, lower = 0, upper = 1)
  assert_number(deltaF, lower = 0, upper = 1)
  assert_number(weights, lower = 1, finite = TRUE)
  assert_number(weightsS, lower = 1, finite = TRUE)
  assert_flag(relativeDelta)

  index_look <- 1
  Nmax <- max(nnr)
  NmaxControl <- max(nnF)
  decision <- NA
  response <- stats::rbinom(Nmax, size = 1, prob = truep)
  # The criteria for Decision 2 for Interim looks are :
  #- Interim GO : P ( successful at final) > phiU
  #- Interim STOP : P ( unsuccessful at final ) > phiFu
  while (is.na(decision) && index_look < length(nnr)) {
    size_look <- nnr[index_look]
    if (size_look %in% nnE) {
      interim_qU <- predprobDist(
        x = sum(response[1:size_look]),
        n = size_look,
        xS = xS,
        nS = nS,
        Nmax = Nmax,
        NmaxControl = NmaxControl,
        delta = deltaE,
        relativeDelta = relativeDelta,
        parE = parE,
        weights = weights,
        parS = parS,
        weightsS = weightsS,
        thetaT = tT
      )$result
      decision <- ifelse(interim_qU > phiU, FALSE, decision)
    }
    if (size_look %in% nnF) {
      interim_qU <- 1 - predprobDist(
        x = sum(response[1:size_look]),
        n = size_look,
        xS = xS,
        nS = nS,
        Nmax = Nmax,
        NmaxControl = NmaxControl,
        delta = deltaF,
        relativeDelta = relativeDelta,
        parE = parE,
        weights = weights,
        parS = parS,
        weightsS = weightsS,
        thetaT = tT
      )$result
      decision <- ifelse(interim_qU > phiFu, FALSE, decision)
    }
    index_look <- index_look + 1
  }
  # The criteria for Decision 2 for Futility looks are :
  #- Final GO = P( RR > p0 + delta ) > tT
  #- Final STOP = P( RR < p1 - delta ) > tF
  if (is.na(decision)) {
    size_look <- nnr[index_look]
    if (size_look %in% nnE) {
      final_eff_qU <- postprobDist(
        x = sum(response[1:size_look]),
        n = size_look,
        xS = xS,
        nS = nS,
        delta = deltaE,
        relativeDelta = relativeDelta,
        parE = parE,
        parS = parS
      )
    }
    decision <- ifelse(final_eff_qU > tT, TRUE, NA)
  }
  if (size_look %in% nnF) {
    final_fu_qU <- 1 - postprobDist(
      x = sum(response[1:size_look]),
      n = size_look,
      xS = xS,
      nS = nS,
      delta = deltaF,
      relativeDelta = relativeDelta,
      parE = parE,
      parS = parS
    )
    decision <- ifelse(final_fu_qU > tF, FALSE, decision)
  }
  list(
    decision = decision,
    all_sizes = size_look
  )
}
```


# Example h_get_decision_two_predprobDis

```{r}
example_dist <- h_get_distance(nn = c(10, 20, 30))
looks <- h_get_looks(dist = example_dist, nnE = c(10, 20, 30), nnF = c(10, 20, 30))
looks_nnrE <- looks$nnrE
looks_nnrF <- looks$nnrF

# input values from above arguments
h_get_decision_two_predprobDist(
  nnE = looks$nnrE,
  nnF = looks$nnrF,
  nnr = c(10, 20, 30),
  truep = 0.2,
  xS = 6,
  nS = 20,
  parE = c(1, 1),
  parS = c(1, 1),
  tT = 0.6,
  tF = 0.6,
  phiU = 0.7,
  phiL = 0.8,
  phiFu = 0.8,
  deltaE = 0.1,
  deltaF = 0.1,
  weights = input$weights,
  weightsS = input$weightsS,
  relativeDelta = TRUE
)
```

# new user facing function

```{r}
new_ocPredprobDist <- function(
    xS = 0,
    nS = 0,
    nnE,
    truep,
    deltaE,
    deltaF,
    relativeDelta = FALSE,
    tT = 1 - tF,
    tF = 1 - tT,
    phiU,
    phiL = 1 - phiFu,
    phiFu = 1 - phiL,
    parE = c(1, 1),
    parS = c(1, 1),
    weights = weights,
    weightsS = weightsS,
    sim = 50000,
    wiggle = FALSE,
    nnF = nnE,
    decision1 = TRUE) {
  Nmax <- max(unique(nnE, nnF))

  assert_number(xS, finite = TRUE)
  assert_number(nS, finite = TRUE)
  assert_numeric(nnE, any.missing = FALSE, sort = TRUE)
  assert_number(truep, lower = 0, upper = 1)
  assert_number(deltaE, lower = 0, upper = 1)
  assert_number(deltaF, lower = 0, upper = 1)
  assert_number(tT, lower = 0, upper = 1)
  assert_number(tF, lower = 0, upper = 1)
  assert_number(phiL, lower = 0, upper = 1)
  assert_number(phiU, lower = 0, upper = 1)
  assert_number(phiFu, lower = 0, upper = 1)
  assert_numeric(parE, min.len = 2, any.missing = FALSE)
  assert_numeric(parS, min.len = 2, any.missing = FALSE)
  assert_number(sim, lower = 1, finite = TRUE)
  assert_flag(wiggle)
  assert_numeric(nnF, lower = 1, any.missing = FALSE, sort = TRUE)
  assert_flag(decision1)

  nn <- sort(unique(c(nnF, nnE)))
  if (sim < 50000) {
    warning("Advise to use sim >= 50000 to achieve convergence")
  }
  decision <- vector(length = sim)
  all_sizes <- vector(length = sim)
  for (k in seq_len(sim)) {
    if (length(nn) != 1 && wiggle) {
      # if we have more than one look in nnF and nnE, we don't wiggle
      dist <- h_get_distance(nn = nn)
      nnr <- h_get_looks(dist = dist, nnE = nnE, nnF = nnF)
      nnrE <- nnr$nnrE
      nnrF <- nnr$nnrF
    } else {
      dist <- 0
      nnrE <- nnE
      nnrF <- nnF
    }
    nnr <- unique(sort(c(nnrE, nnrF)))
    tmp <- if (decision1) {
      h_get_decision_one_predprobDist(
        nnE = nnE,
        nnF = nnF,
        nnr = nnr,
        truep = truep,
        xS = xS,
        nS = nS,
        parE = parE,
        parS = parS,
        tT = tT,
        phiU = phiU,
        phiL = phiL,
        deltaE = deltaE,
        deltaF = deltaF,
        weights = weights,
        weightsS = weightsS,
        relativeDelta = relativeDelta
      )
    } else {
      h_get_decision_two_predprobDist(
        nnE = nnE,
        nnF = nnF,
        nnr = nnr,
        truep = truep,
        xS = xS,
        nS = nS,
        parE = parE,
        parS = parS,
        tT = tT,
        tF = tF,
        phiU = phiU,
        phiL = phiL,
        phiFu = phiFu,
        deltaE = deltaE,
        deltaF = deltaF,
        weights = weights,
        weightsS = weightsS,
        relativeDelta = relativeDelta
      )
    }
    decision[k] <- tmp$decision
    all_sizes[k] <- tmp$all_sizes
  }
  oc <- h_get_oc_predprob(all_sizes = all_sizes, nnr = nnr, decision = decision)
  list(
    oc = oc,
    Decision = decision,
    SampleSize = all_sizes,
    wiggled_nnrE = nnrE,
    wiggled_nnrF = nnrF,
    dist = dist,
    params = as.list(match.call(expand.dots = FALSE))
  )
}
```

# Examples to compare with vintage function
```{r}
set.seed(20)
res1 <- ocPredprobDist( # this is equivalent to decision 1, and relativeDelta = FALSE
  nn = c(10, 20, 30),
  p = 0.4,
  delta = 0.1,
  tT = 0.6,
  phiL = 0.2,
  phiU = 0.8,
  parE = c(1, 1),
  parS = c(5, 25),
  ns = 10
)
res1$oc
```
# Example for new function
```{r}
set.seed(20)
new_ocPredprobDist(
  xS = 0,
  nS = 0,
  nnE = c(10, 20, 30),
  truep = 0.4,
  deltaE = 0.1,
  deltaF = 0.1,
  relativeDelta = FALSE,
  tT = 0.6,
  phiU = 0.8,
  phiL = 0.2,
  # phiFu = 1 - phiL,
  parE = c(1, 1),
  parS = c(5, 25),
  weights = input$weights,
  weightsS = input$weightsS,
  sim = 10,
  wiggle = FALSE,
  nnF = c(10, 20, 30),
  decision1 = TRUE
)
```


```{r}
res4 <- ocPredprobDist(c(10, 20, 30), 0.4,
  delta = 0.1, tT = 0.6, phiL = 0.2, phiU = 0.8,
  parE = c(1, 1), parS = c(5, 25), ns = 10, nr = TRUE, d = 5, nnF = c(30)
)
res4$oc
```
