---
title: "Design Document for ocPredprob.R"
output: html_document
date: "2024-01-08"
editor_options:
  chunk_output_type: console
---
## Rule 1
"Decision rule 1:The trial is stopped for efficacy if the predictive probability of a
successful trial is larger than phiU, and stopped for futility if it is below
phiL. A trial is successful if after the maximum number of patients the
posterior probability of the treatment having more than p0 response rate is
above tT. Otherwise the decision is "failure". In this case, there is no gray zone."


Translation:
* Decision 1 says:
    * Final GO = P(p > p0 | data) => tT, by the way P(truep > p0) > tT = qU
    when the prob of good response is high enough, then we have a successful trial
    * Final STOP = P(p > p0 | data ) < tT =  Unsuccessful trial if not successful
    * Final EVAL = there is no evaluate
    * interim GO =  P(successful trial at final) > phiU, P(interim_qU > tT) > phiU
    * interim STOP = P(successful trial at final) < phiL, P(interim_qU > tT) < phiL
    * interim EVAL = in between

## Rule 2
Decision rule 2:A variation can be requested when skipping the argument phiL and utilizing the arguments
p1, tFu & PhiFu. The trial can be stopped for futility if the predictive
probability of an unsuccessful trial is larger than phiFu. In this case, the
decision is "failure" when the posterior probability of having the treatment
response rate at most p1 is above tFu at the maximum number of patients.

Translation:
* Decision 2 says:
    * Final GO = P(truep > p0) > tT =  when the prob of good response is high, then we have a successful trial
    * Final STOP = P(truep < p1) > tF = when the prob of poor response is high, then we have an unsuccessful trial
    * Final EVAL = in between
    * interim STOP = P (failure at final ) > phiFu
    * interim GO = P ( success at final) > phiU
    * interim EVAL = in between

## Design notes :

* both decisions converge with old code

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1 Helper function for Decision
    * Decision 1 says:
    * GO = P(truep > p0 | data) => tT, by the way P(truep > p0) > tT = qU
    when the prob of good response is high enough, then we have a successful trial
    * STOP = P(truep > p0 | data ) < tT =  Unsuccessful trial if not successful
    * EVAL = there is no evaluate BUT there are interims.... in this code
    * interim GO =  P(successful trial at final) > phiU, P(interim_qU > tT) > phiU
    * interim STOP = P(successful trial at final) < phiL, P(interim_qU > tT) < phiL
    * interim EVAL = in between
```{r, echo = TRUE}
h_get_decision_one_predprob <- function(nnr, truep, p0, parE = c(1, 1), nnE, nnF, tT, phiU, phiL) {
  index_look <- 1
  assert_numeric(nnr)
  all_sizes <- decision <- NA
  response <- stats::rbinom(max(nnr), size = 1, truep)
  assert_numeric(response, lower = 0, upper = 1)
  while (is.na(decision) && index_look < length(nnr)) { # at interim
    size_look <- nnr[index_look]
    if (size_look %in% nnE) { # GO decision for interim
      interim_qU <- predprob(
        x = sum(x = response[1:size_look]),
        n = size_look,
        Nmax = nnE[length(nnE)],
        p = p0, # p1 only used if decision2 is specified so it does not exist here
        thetaT = tT,
        parE = parE
      )$result
      decision <- ifelse(interim_qU > phiU, TRUE, NA)
      assert_flag(decision, na.ok = TRUE)
    }
    if (size_look %in% nnF) {
      interim_qU <- predprob( # STOP decision for interim
        x = sum(x = response[1:size_look]),
        n = size_look,
        Nmax = nnF[length(nnF)],
        p = p0,
        thetaT = tT,
        parE = parE
      )$result
      decision <- ifelse(interim_qU < phiL, FALSE, decision)
      assert_flag(decision, na.ok = TRUE)
    }
    index_look <- index_look + 1
  }
  if (is.na(decision)) {
    size_look <- nnr[index_look]
    # at final
    if (size_look %in% nnE) { # for efficacy looks
      final_eff_qU <- postprob(
        x = sum(x = response[1:size_look]),
        n = size_look,
        p = p0,
        parE = parE,
        log.p = FALSE
      )
      # it is an final GO if P(response rate for GO > tT ) > phiU
      decision <- ifelse(final_eff_qU > tT, TRUE, NA)
      assert_flag(decision, na.ok = TRUE)
    }
    if (size_look %in% nnF) { # for futility looks
      final_fu_qU <- postprob(
        x = sum(x = response[1:size_look]),
        n = size_look,
        p = p0,
        parE = parE,
        log.p = FALSE
      )
      # it is an final STOP if P(response rate for STOP < tT )  < phiL
      decision <- ifelse(final_fu_qU < tT, FALSE, decision)
      assert_flag(decision, na.ok = TRUE)
    }
  }
  list(
    decision = decision,
    all_sizes = size_look
  )
}
```
# Example for decision 1
if nnE = c(10, 20) and nnF =  c(11, 15), and the rules are:

Final:
GO: P (truep > p0 = 0.2) > tT = 0.6
STOP = P(truep > p0 = 0.2) < tT = 0.6 # no grey zone

Interim:
GO: P (successful trial) > phiU = 0.5
STOP: P (successful trial) < phiL = 0.5 # no grey zone

Should be a Go since the truep is assumed to be 0.6 and the threshold for a good response (p0) is 0.2, and the final threshold is only 0.6.

```{r}
h_get_decision_one_predprob(
  nnr = c(10, 20, 30),
  truep = 0.4,
  p0 = 0.25,
  parE = c(1, 1),
  nnE = c(10, 20, 30),
  nnF = c(10, 20, 30),
  tT = 0.6,
  phiU = 0.8,
  phiL = 0.3
)
# (nnr, truep, p0, parE = c(1, 1), nnE, nnF, tT, phiU, phiL)
```

## 2 Helper function for Decision 2
    * Decision 2 says:
    * Final GO = P(truep > p0) > tT =  when the prob of good response is high, then we have a successful trial
    * Final STOP = P(truep < p1) > tF = when the prob of poor response is high, then we have an unsuccessful trial
    * interim STOP : P (failure at final ) > phiFu
    * interim GO : P ( success at final) > phiU
```{r}
h_get_decision_two_predprob <- function(nnr, truep, p0, p1, parE = c(1, 1), nnE, nnF, tT, tF, phiFu, phiU) {
  index_look <- 1
  assert_numeric(nnr)
  all_sizes <- decision <- NA
  response <- stats::rbinom(max(nnr), size = 1, truep)
  assert_numeric(response, lower = 0, upper = 1)
  while (is.na(decision) && index_look < length(nnr)) { # as long as there is no decision...
    size_look <- nnr[index_look]
    if (size_look %in% nnE) {
      interim_qU <- predprob(
        x = sum(x = response[1:size_look]),
        n = size_look,
        Nmax = nnE[length(nnE)],
        p = p0,
        thetaT = tT,
        parE = parE
      )$result
      decision <- ifelse(interim_qU > phiU, TRUE, NA)
    }
    if (size_look %in% nnF) {
      interim_qU <- 1 - predprob(
        x = sum(x = response[1:size_look]),
        n = size_look,
        Nmax = nnF[length(nnF)],
        p = p1,
        thetaT = tF,
        parE = parE
      )$result
      decision <- ifelse(interim_qU > phiFu, FALSE, decision)
    }
    index_look <- index_look + 1
  }
  if (is.na(decision)) {
    if (size_look %in% nnE) { # for efficacy looks at FINAL
      final_qU <- postprob(
        # based on all data, the posterior probability is a GO when P(truep > p0) > tT
        x = response,
        n = size_look,
        p = truep,
        parE = c(1, 1),
        log.p = FALSE
      )
      decision <- ifelse(final_qU > tT, TRUE, NA)
    }
    if (size_look %in% nnF) { # for futility looks at FINAL
      # based on all data, the posterior probability is a STOP when P(truep > p0) > tF
      final_qU <- 1 - postprob(
        x = response,
        n = size_look,
        p = truep,
        parE = c(1, 1),
        log.p = FALSE
      )
      decision <- ifelse(final_qU > tF, FALSE, decision)
    }
  }
  list(
    decision = decision,
    all_sizes = size_look
  )
}
```
# Example for decision 2
if nnE = c(10, 20) and nnF =  c(10, 20), and the rules are:

Final:
GO: P (truep > p0 = 0.2) > tT = 0.6
STOP = P(truep < p1 = 0.2) > tF = 0.6 # no grey zone

Interim:
GO: P (successful trial) > phiU = 0.5
STOP: P (successful trial) > phiFu = 0.5 # no grey zone

when truep = 0.4 and the threshold is lower at 0.2 = p0 = p1.
```{r}
h_get_decision_two_predprob(
  nnr = c(10, 20),
  truep = 0.4,
  p0 = 0.2,
  p1 = 0.2,
  parE = c(1, 1),
  nnE = c(10, 20),
  nnF = c(10, 20),
  tT = 0.6,
  tF = 0.6,
  phiFu = 0.5,
  phiU = 0.5
)

# (nnr, truep, p0, p1, parE = c(1, 1), nnE, nnF, tT, tF, phiFu, phiU)
```

# helper h_get_oc_predprob : the last helper step in the user interfacing function
```{r}
h_get_oc_predprob <- function(all_sizes, nnr, decision, nnrE, nnrF, Nmax = Nmax) {
  sim <- length(all_sizes)
  # assert_logical(decision, len = sim)
  # assert_numeric(all_sizes)
  # assert_numeric(nnrE, lower = 0, upper = max(nnrF))
  # assert_numeric(nnrF, lower = 0, upper = max(nnrF))



  assert_numeric(all_sizes, any.missing = FALSE)
  assert_numeric(nnr, lower = 1, upper = Nmax)
  assert_logical(decision)
  assert_numeric(nnrE, lower = 1, upper = Nmax)
  assert_numeric(nnrE, lower = 1, upper = Nmax)
  assert_number(Nmax, upper = Nmax)
  assert_number(sim, lower = 1)

  data.frame(
    ExpectedN = mean(all_sizes, na.rm = TRUE),
    PrStopEarly = mean(all_sizes < max(nnrF), na.rm = TRUE),
    PrEarlyEff = sum(decision * (all_sizes < max(nnrE)), na.rm = TRUE) / sim,
    PrEarlyFut = sum((1 - decision) * (all_sizes < max(nnrF)), na.rm = TRUE) / sim,
    PrEfficacy = sum(decision, na.rm = TRUE) / sim,
    PrFutility = sum(1 - decision, na.rm = TRUE) / sim,
    PrGrayZone = sum(is.na(decision) / sim)
  )
}
```

## new_ocPredprob

```{r cars}
new_ocPredprob <- function(nnE, truep, p0, p1, tT, tF, phiL, phiU, phiFu, parE = c(1, 1),
                           sim = 50000, wiggle = FALSE, nnF = nnE, decision1 = TRUE) {
  nn <- sort(unique(c(nnF, nnE)))
  assert_number(sim, lower = 1, finite = TRUE)
  assert_flag(wiggle)
  assert_flag(decision1)
  if (sim < 50000) {
    warning("Advise to use sim >= 50000 to achieve convergence")
  }
  decision <- vector(length = sim)
  all_sizes <- vector(length = sim)

  for (k in seq_len(sim)) {
    if (length(nn) != 1 && wiggle) {
      dist <- h_get_distance(nn = nn)
      nnr <- h_get_looks(dist = dist, nnE = nnE, nnF = nnF)
      nnrE <- nnr$nnrE
      nnrF <- nnr$nnrF
    } else {
      dist <- 0
      nnrE <- nnE
      nnrF <- nnF
    }
    nnr <- unique(c(nnrE, nnrF))
    # Nmax <- max(nnr)
    tmp <- if (decision1) {
      h_get_decision_one_predprob(
        nnr = nnr,
        truep = truep,
        p0 = p0,
        parE = parE,
        nnE = nnrE,
        nnF = nnrF,
        tT = tT,
        phiU = phiU,
        phiL = phiL
      )
    } else {
      h_get_decision_two_predprob(
        nnr = nnr,
        truep = truep,
        p0 = p0,
        p1 = p1,
        parE = parE,
        nnE = nnrE,
        nnF = nnrF,
        tT = tT,
        tF = tF,
        phiFu = phiFu,
        phiU = phiU
      )
    }
    decision[k] <- tmp$decision
    all_sizes[k] <- tmp$all_sizes
  }
  oc <- h_get_oc_predprob(all_sizes = all_sizes, nnr = nnr, decision = decision, nnrE = nnrE, nnrF = nnrF)
  list(
    oc = oc,
    Decision = decision,
    SampleSize = all_sizes,
    union_nn = nnr,
    input_nnE = nnE,
    input_nnF = nnF,
    wiggled_nnE = nnrE,
    wiggled_nnF = nnrF,
    wiggle_dist = dist,
    params = as.list(match.call(expand.dots = FALSE))
  )
}
```


```{r}
set.seed(100)
new_ocPredprob(
  nnE = c(10, 20, 30),
  truep = 0.4,
  p0 = 0.25,
  # p1 = 0.30,
  tT = 0.6,
  # tF = 0.5,
  phiL = 0.3,
  phiU = 0.8,
  parE = c(1, 1),
  sim = 10,
  wiggle = FALSE,
  nnF = c(10, 20, 30),
  decision1 = TRUE
)
```

# Are we getting the same results ? We simulate sim number of trials to check if in decision 1 and decision 2, we are getting the same results
1. Run ten simulations of decision 1, compare it with the grandma function
2. Run ten simulations of decision 2, compare it with the grandma function

# testing dec 1 ten times
```{r}
set.seed(10)
new_ocPredprob(
  nnE = c(10, 20, 30),
  truep = 0.4,
  p0 = 0.25,
  # p1 = 0.30,
  tT = 0.6,
  # tF = 0.5,
  phiL = 0.3,
  phiU = 0.8,
  parE = c(1, 1),
  sim = 10,
  wiggle = FALSE,
  nnF = c(10, 20, 30),
  decision1 = TRUE
)
# $oc
#   ExpectedN PrStopEarly PrEarlyEff PrEarlyFut PrEfficacy PrFutility PrGrayZone
# 1        15           1        0.8        0.2        0.8        0.2          0
ocPredprob( # no wiggle is default
  nn = c(10, 20, 30),
  p = 0.4, # truep
  p0 = 0.25,
  tT = 0.6,
  phiL = 0.3,
  phiU = 0.8,
  parE = c(1, 1),
  ns = 10
)
# $oc
#      ExpectedN PrStopEarly PrEarlyEff PrEarlyFut PrEfficacy PrFutility PrGrayZone
# [1,]        13         0.9        0.9          0          1          0          0
# ---------------------
# more simulations
set.seed(10)
result <- new_ocPredprob(
  nnE = c(10, 20, 30),
  truep = 0.4,
  p0 = 0.25,
  p1 = 0.30,
  tT = 0.6,
  # tF = 0.5,
  phiL = 0.3,
  phiU = 0.8,
  parE = c(1, 1),
  sim = 1000,
  wiggle = FALSE,
  nnF = c(10, 20, 30),
  decision1 = TRUE
)
result$oc
# for 1000 sims
#   ExpectedN PrStopEarly PrEarlyEff PrEarlyFut PrEfficacy PrFutility PrGrayZone
# 1     14.08       0.937      0.858      0.079      0.858      0.142          0
expected <- ocPredprob(
  nn = c(10, 20, 30),
  p = 0.4, # truep
  p0 = 0.25,
  tT = 0.6,
  phiL = 0.3,
  phiU = 0.8,
  parE = c(1, 1),
  ns = 1000
)
expected$oc
table(result$SampleSize)
table(expected$SampleSize)
# $oc
#      ExpectedN PrStopEarly PrEarlyEff PrEarlyFut PrEfficacy PrFutility PrGrayZone
# [1,]      13.8       0.959      0.886      0.073      0.921      0.079          0

# ---------------------
# more simulations
set.seed(10)
result <- new_ocPredprob(
  nnE = c(10, 20, 30),
  truep = 0.4,
  p0 = 0.25,
  # p1 = 0.30,
  tT = 0.6,
  # tF = 0.5,
  phiL = 0.3,
  phiU = 0.8,
  parE = c(1, 1),
  sim = 10000,
  wiggle = FALSE,
  nnF = c(10, 20, 30),
  decision1 = TRUE
)
# $oc
#   ExpectedN PrStopEarly PrEarlyEff PrEarlyFut PrEfficacy PrFutility PrGrayZone
# 1    13.938      0.9476     0.8713     0.0763     0.8713     0.1287          0
expected <- ocPredprob(
  nn = c(10, 20, 30),
  p = 0.4, # truep
  p0 = 0.25,
  tT = 0.6,
  phiL = 0.3,
  phiU = 0.8,
  parE = c(1, 1),
  ns = 10000
)$oc
table(result$SampleSize)
table(expected$SampleSize)
# $oc
#      ExpectedN PrStopEarly PrEarlyEff PrEarlyFut PrEfficacy PrFutility PrGrayZone
# [1,]    13.862      0.9501     0.8725     0.0776     0.9142     0.0858          0




#--- we changed it to sum(x) in responses:
set.seed(10)
new_ocPredprob(
  nnE = 30,
  truep = 0.4,
  p0 = 0.25,
  # p1 = 0.30,
  tT = 0.6,
  # tF = 0.5,
  phiL = 0.3,
  phiU = 0.8,
  parE = c(1, 1),
  sim = 10000,
  wiggle = FALSE,
  nnF = c(10, 20, 30),
  decision1 = TRUE
)$oc
# ExpectedN PrStopEarly PrEarlyEff PrEarlyFut PrEfficacy PrFutility PrGrayZone
# 1        30           0          0          0      0.954      0.046          0
ocPredprob( # no wiggle is default
  nn = 30,
  p = 0.4, # truep
  p0 = 0.25,
  tT = 0.6,
  phiL = 0.3,
  phiU = 0.8,
  parE = c(1, 1),
  ns = 10000
)$oc
# ExpectedN PrStopEarly PrEarlyEff PrEarlyFut PrEfficacy PrFutility PrGrayZone
# [1,]        30           0          0          0     0.9583     0.0417          0


new_ocPredprob(
  nnE = 30,
  truep = 0.4,
  p0 = 0.25,
  # p1 = 0.30,
  tT = 0.6,
  # tF = 0.5,
  phiL = 0.3,
  phiU = 0.8,
  parE = c(1, 1),
  sim = 50000,
  wiggle = FALSE,
  nnF = c(10, 20, 30),
  decision1 = TRUE
)$oc
# ExpectedN PrStopEarly PrEarlyEff PrEarlyFut PrEfficacy PrFutility PrGrayZone
# 1        30           0          0          0    0.95678    0.04322          0
ocPredprob( # no wiggle is default
  nn = 30,
  p = 0.4, # truep
  p0 = 0.25,
  tT = 0.6,
  phiL = 0.3,
  phiU = 0.8,
  parE = c(1, 1),
  ns = 50000
)$oc
# + )$oc
# ExpectedN PrStopEarly PrEarlyEff PrEarlyFut PrEfficacy PrFutility PrGrayZone
# [1,]        30           0          0          0    0.95542    0.04458          0
```

# testing dec 2 ten times

```{r}
set.seed(10)
new_ocPredprob(
  nnE = c(10, 20, 30),
  truep = 0.4,
  p0 = 0.25,
  p1 = 0.30,
  tT = 0.6,
  tF = 0.5,
  phiL = NA,
  phiU = 0.8,
  phiFu = 0.5,
  parE = c(1, 1),
  sim = 10,
  wiggle = FALSE,
  nnF = c(10, 20, 30),
  decision1 = FALSE
)
# $oc
#   ExpectedN PrStopEarly PrEarlyEff PrEarlyFut PrEfficacy PrFutility PrGrayZone
# 1        12           1        0.7        0.3        0.7        0.3          0
ocPredprob(
  nn = c(10, 20, 30),
  p = 0.4, # truep
  p0 = 0.25,
  p1 = 0.30,
  tT = 0.6,
  tF = 0.5,
  # phiL = NA,
  phiU = 0.8,
  phiFu = 0.5,
  parE = c(1, 1),
  nr = FALSE,
  d = NULL,
  ns = 10
)
# $oc
#      ExpectedN PrStopEarly PrEarlyEff PrEarlyFut PrEfficacy PrFutility PrGrayZone
# [1,]        11           1        0.8        0.2        0.8        0.2          0

#--- more iterations---
set.seed(10)
result <- new_ocPredprob(
  nnE = c(10, 20, 30),
  truep = 0.4,
  p0 = 0.25,
  p1 = 0.30,
  tT = 0.6,
  tF = 0.5,
  phiL = NA,
  phiU = 0.8,
  phiFu = 0.5,
  parE = c(1, 1),
  sim = 100,
  wiggle = FALSE,
  nnF = c(10, 20, 30),
  decision1 = FALSE
)

result$oc
# > result$oc
#   ExpectedN PrStopEarly PrEarlyEff PrEarlyFut PrEfficacy PrFutility PrGrayZone
# 1      11.6           1        0.8        0.2        0.8        0.2          0
expected <- ocPredprob(
  nn = c(10, 20, 30),
  p = 0.4, # truep
  p0 = 0.25,
  p1 = 0.30,
  tT = 0.6,
  tF = 0.5,
  # phiL = NA,
  phiU = 0.8,
  phiFu = 0.5,
  parE = c(1, 1),
  nr = FALSE,
  d = NULL,
  ns = 100
)
expected$oc
#      ExpectedN PrStopEarly PrEarlyEff PrEarlyFut PrEfficacy PrFutility PrGrayZone
# [1,]      11.9           1       0.76       0.24       0.76       0.24          0
table(result$SampleSize)
table(expected$SampleSize)

#--- more and more iterations---
set.seed(10)
result <- new_ocPredprob(
  nnE = c(10, 20, 30),
  truep = 0.4,
  p0 = 0.25,
  p1 = 0.30,
  tT = 0.6,
  tF = 0.5,
  phiL = NA,
  phiU = 0.8,
  phiFu = 0.5,
  parE = c(1, 1),
  sim = 10000,
  wiggle = FALSE,
  nnF = c(10, 20, 30),
  decision1 = FALSE
)

result$oc
# > result$oc
#   ExpectedN PrStopEarly PrEarlyEff PrEarlyFut PrEfficacy PrFutility PrGrayZone
# 1     12.19           1     0.7975     0.2025     0.7975     0.2025          0

expected <- ocPredprob(
  nn = c(10, 20, 30),
  p = 0.4, # truep
  p0 = 0.25,
  p1 = 0.30,
  tT = 0.6,
  tF = 0.5,
  # phiL = NA,
  phiU = 0.8,
  phiFu = 0.5,
  parE = c(1, 1),
  nr = FALSE,
  d = NULL,
  ns = 10000
)
expected$oc
# > expected$oc
#      ExpectedN PrStopEarly PrEarlyEff PrEarlyFut PrEfficacy PrFutility PrGrayZone
# [1,]    12.118           1     0.7926     0.2074     0.7926     0.2074          0

table(result$SampleSize)
table(expected$SampleSize)
```

# Creating global variables for final results testing
```{r, echo = FALSE}
set.seed(2024)
tt <- list(
  nnE = c(10, 20, 30), nnF = c(10, 20, 30), truep = 0.4, p0 = 0.25, p1 = 0.30, tT = 0.6, tF = 0.5, tFu = 0.6,
  phiU = 0.8, phiL = 0.3, phiFu = 0.5, sim = 100, decision1 = TRUE, parE = c(1, 1), size_look = 30,
  wiggle = TRUE, sim = 500
)

nnE <- tt$nnE
nnF <- sort(tt$nnF)
s <- rep(NA, tt$sim)
truep <- tt$truep
n <- s
nn <- sort(unique(c(tt$nnF, tt$nnE)))
nL <- length(nn)
Nstart <- nn[1]
Nmax <- nn[nL]
decision1 <- tt$decision1
# d = FALSE
p <- tt$truep
p1 <- tt$p0
p0 <- tt$p0
tFu <- 1 - tt$tT
parE <- tt$parE
tT <- tt$tT
ns <- tt$sim
nr <- FALSE
phiFu <- tt$phiFu
phiU <- tt$phiU
phiL <- tt$phiL
size_look <- tt$size_look
wiggle <- tt$wiggle
sim <- tt$sim
```
