---
title: "Design document for ocRctPostprobDist"
output: html_notebook
---

Calculate operating characteristics for RCT against SOC,
using the posterior probability method with beta priors

We emulate a randomized-controlled trial setting where at any given sample size,
there exists the number of patients enrolled in either standard of care (SOC) or control arm, and
a treatment or experimental arm. The allocation of patients will depend on the
randomization ratio set by the user and is rounded to the next higher integer.
Therefore the sequence of patients is determined from the start, such that the number of
patients in both arms is constant across trial simulations, however the number of patients
within the control and treatment arm is determined by the randomisation ratio.
Interim looks are for sample sizes below that of the final sample size.
Final looks are only performed at the maximum sample size,
At each interim or final look, a futility or efficacy or both can be performed.

The rules for Stop, Go and Gray Zone (where applicable) are the same as in [ocPostprobDist()] where
the only difference is to emulate a randomized-controlled trial setting.

The returned value is a list with the following elements:
 - `oc`: matrix with operating characteristics with the following details:
    - `ExpectedN`: expected number of patients in the trials in both treatment and SOC group
    - `ExpectedNactive` : the mean of the number of patients in treatment arm
    - `ExpectedNcontrol`: the mean of the number of patients in control arm
    - `PrStopEarly`: probability to stop the trial early (before reaching the
                    maximum sample size)
    - `PrEarlyEff`: probability of Early Go decision
    - `PrEarlyFut`: probability of Early Stop decision
    - `PrEfficacy`: probability of Go decision
    - `PrFutility`: probability of Stop decision
    - `PrGrayZone`: probability of Evaluate or "Gray Zone" decision (between Go and Stop)
 - `Decision` : numeric of results with `TRUE` as Go, `FALSE` as Stop and `NA` as Evaluate decision.
 - `SampleSize` : numeric of sample sizes from `nnE` or `nnF` or both.
 - `wiggled_nnE` : user input for `nnE` with random distance applied.
 - `wiggled_nnF` : user input for `nnF` with random distance applied.
 - `wiggled_dist` : magnitude of random distance applied in order of input looks.
 - `params` : all user input arguments.

# Input parameters
```{r}
input <- list(
  nn = c(10, 20, 30),
  nnE = c(10, 20, 30),
  pE = 0.4,
  pS = 0.3,
  deltaE = 0.2,
  deltaF = 0.1,
  relativeDelta = FALSE,
  tL = 0.8,
  tU = 0.8,
  parE = c(a = 1, b = 1),
  parS = c(a = 1, b = 1),
  randRatio = 1,
  ns = 10000,
  nr = FALSE,
  d = NULL,
  nnF = c(10, 20, 30),
  truep = 0.4,
  sim = 5,
  Nmax = 15,
  NmaxControl = 15,
  decision <- rep(NA, 5)
)
n <- nActive <- nControl <- integer(5) # initiate n
ns <- as.integer(5) # sim
nL <- length(input$nn) # the max of nn or we can do Nmax
Nstart <- input$nn[1]
Nmax <- input$nn[nL]
```
# Example of vintage function
```{r}
ocRctPostprobDist(
  nn = input$nn,
  pE = input$pE,
  pS = input$pS,
  deltaE = input$deltaE,
  deltaF = input$deltaF,
  relativeDelta = input$relativeDelta,
  tL = input$tL,
  tU = input$tU,
  parE = input$parE,
  parS = input$parS,
  randRatio = 1,
  ns = input$sim,
  nr = input$nr,
  d = input$d,
  nnF = input$nnF
)
```
# h_get_decisionDist_rct
```{r}
h_get_decisionDist_rct <- function(nnr,
                                   nnrE,
                                   nnrF,
                                   pE,
                                   pS,
                                   parE = c(1, 1),
                                   parS = c(1, 1),
                                   tL,
                                   tU,
                                   deltaE,
                                   deltaF,
                                   relativeDelta,
                                   randRatio = 1) {
  assert_numeric(nnr, finite = TRUE, any.missing = FALSE)
  assert_numeric(nnrE, max.len = length(nnr), any.missing = FALSE)
  assert_numeric(nnrF, max.len = length(nnr), any.missing = FALSE)
  assert_number(pE, lower = 0, upper = 1)
  assert_number(pS, lower = 0, upper = 1)
  assert_numeric(parE, lower = 0, finite = TRUE, any.missing = FALSE)
  assert_numeric(parS, lower = 0, finite = TRUE, any.missing = FALSE)
  assert_number(tL, lower = 0, upper = 1)
  assert_number(tU, lower = 0, upper = 1)
  assert_number(deltaE, finite = TRUE)
  assert_number(deltaF, finite = TRUE)
  assert_flag(relativeDelta)
  assert_number(randRatio, na.ok = FALSE, upper = 1, finite = TRUE)

  index_look <- 1
  size_look <- nnr[index_look]
  all_sizes <- decision <- nActive <- nControl <- NA
  activeProp <- randRatio / (randRatio + 1)
  NmaxActive <- ceiling(activeProp * Nmax)
  NmaxControl <- Nmax - NmaxActive

  isActive <- sample(
    x = rep(c(TRUE, FALSE), c(NmaxActive, NmaxControl)),
    size = Nmax,
    replace = FALSE
  )
  response <- rbinom(Nmax, size = 1, prob = ifelse(isActive, pE, pS))

  activeProp <- randRatio / (randRatio + 1)
  NmaxActive <- ceiling(activeProp * Nmax)
  NmaxControl <- Nmax - NmaxActive


  while (is.na(decision) && index_look <= length(nnr)) {
    ## current data in both arms:
    xActive <- response[which(isActive[1:size_look])]
    xControl <- response[which(!isActive[1:size_look])]

    if (size_look %in% nnrF) {
      qL <- postprobDist(
        x = sum(xControl),
        n = length(xControl),
        xS = sum(xActive),
        nS = length(xActive),
        delta = deltaF,
        relativeDelta = relativeDelta,
        parE = parS,
        parS = parE
      )
      decision <- ifelse(qL >= tL, FALSE, NA)
    }
    if (size_look %in% nnrE) {
      qU <- postprobDist(
        x = sum(xActive),
        n = length(xActive),
        xS = sum(xControl),
        nS = length(xControl),
        delta = deltaE,
        relativeDelta = relativeDelta,
        parE = parE,
        parS = parS
      )
      decision <- ifelse(qU < tU, decision, TRUE)
    }
    nActive <- length(xActive)
    nControl <- length(xControl)
    all_sizes <- size_look
    index_look <- index_look + 1
    size_look <- nnr[index_look]
  }
  list(
    decision = decision,
    all_sizes = all_sizes,
    nActive = nActive,
    nControl = nControl
  )
}
```
# Example of running `h_get_decisionDist_rct`
```{r}
example_dist <- h_get_distance(nn = c(10, 20, 30))
looks <- h_get_looks(dist = example_dist, nnE = c(10, 20, 30), nnF = c(10, 20, 30))
looks_nnrE <- looks$nnrE
looks_nnrF <- looks$nnrF

h_get_decisionDist_rct(
  nnr = input$nn,
  nnrE = looks$nnrE,
  nnrF = looks$nnrF,
  pE = input$pE,
  pS = input$pS,
  parE = input$parE,
  parS = input$parS,
  tL = input$tL,
  tU = input$tU,
  deltaE = input$deltaE,
  deltaF = input$deltaF,
  relativeDelta = TRUE
)
```
# helper also works when `relativeDelta = FALSE`
```{r}
example_dist <- h_get_distance(nn = c(10, 20, 30))
looks <- h_get_looks(dist = example_dist, nnE = c(10, 20, 30), nnF = c(10, 20, 30))
looks_nnrE <- looks$nnrE
looks_nnrF <- looks$nnrF

h_get_decisionDist_rct(
  nnr = input$nn,
  nnrE = looks$nnrE,
  nnrF = looks$nnrF,
  pE = input$pE,
  pS = input$pS,
  parE = c(1, 1),
  parS = c(1, 1),
  tL = input$tL,
  tU = input$tU,
  deltaE = input$deltaE,
  deltaF = input$deltaF,
  relativeDelta = FALSE
)
```
# h_get_oc_rct
```{r}
h_get_oc_rct <- function(all_sizes, nnr, Nmax, nActive, nControl, decision) {
  assert_numeric(all_sizes, any.missing = FALSE)
  assert_numeric(nnr, lower = 1)
  assert_number(Nmax, lower = 1)
  assert_numeric(nActive, any.missing = FALSE)
  assert_numeric(nControl, any.missing = FALSE)
  assert_logical(decision, len = length(all_sizes))
  assert_true(length(all_sizes) == length(decision))
  sim <- length(all_sizes)

  tmp <- h_get_oc_predprob(
    all_sizes = all_sizes,
    nnr = nnr,
    decision = decision
  )
  data.frame(
    ExpectedN = tmp$ExpectedN,
    ExpectedNactive = mean(nActive),
    ExpectedNcontrol = mean(nControl),
    PrStopEarly = tmp$PrStopEarly,
    PrEarlyEff = tmp$PrEarlyEff,
    PrEarlyFut = tmp$PrEarlyFut,
    PrEfficacy = tmp$PrEfficacy,
    PrFutility = tmp$PrFutility,
    PrGrayZone = tmp$PrGrayZone
  )
}
```
# Example for h_get_oc_rct
```{r}
set.seed(2000)
result <- ocRctPostprobDist(
  nn = input$nn,
  pE = input$pE,
  pS = input$pS,
  deltaE = input$deltaE,
  deltaF = input$deltaF,
  relativeDelta = TRUE,
  tL = input$tL,
  tU = input$tU,
  parE = input$parE,
  parS = input$parS,
  randRatio = 1,
  ns = 100,
  nr = input$nr,
  d = input$d,
  nnF = input$nnF
)

oc <- h_get_oc_rct(
  all_sizes = result$SampleSize,
  nnr = input$nn,
  Nmax = max(input$nn),
  nActive = result$SampleSizeActive,
  nControl = result$SampleSizeControl,
  decision = result$Decision
)
oc
```
# User facing function
```{r}
new_RctPostprobDist <- function(nnE,
                                pE, # truep
                                pS,
                                deltaE,
                                deltaF,
                                relativeDelta = FALSE,
                                tL,
                                tU,
                                parE = c(a = 1, b = 1),
                                parS = c(a = 1, b = 1),
                                randRatio = 1,
                                sim,
                                wiggle = FALSE,
                                nnF = nnE) {
  # some starting inits ocPostprobDist
  if (sim < 50000) {
    warning("Advise to use sim >= 50000 to achieve convergence")
  }
  decision <- all_sizes <- vector(length = sim)
  nnE <- sort(nnE)
  nnF <- sort(nnF)
  nn <- sort(unique(c(nnF, nnE)))
  nL <- length(nn)
  nn <- sort(unique(c(nnF, nnE)))

  nL <- length(nn)
  Nstart <- nn[1]
  Nmax <- nn[nL]

  for (k in seq_len(sim)) {
    if (length(nn) != 1 && wiggle) {
      dist <- h_get_distance(nn = nn)
      nnr <- h_get_looks(dist = dist, nnE = nnE, nnF = nnF)
      nnrE <- nnr$nnrE
      nnrF <- nnr$nnrF
    } else {
      nnrE <- nnE
      nnrF <- nnF
    }
    nnr <- unique(c(nnrE, nnrF))
    tmp <- h_get_decisionDist_rct(
      nnr = nnr,
      nnrE = nnrE,
      nnrF = nnrF,
      pE = pE,
      pS = pS,
      parE = c(1, 1),
      parS = c(1, 1),
      tL = tL,
      tU = tU,
      deltaE = deltaE,
      deltaF = deltaF,
      relativeDelta = relativeDelta
    )
    decision[k] <- tmp$decision
    all_sizes[k] <- tmp$all_sizes
    nActive[k] <- tmp$nActive
    nControl[k] <- tmp$nControl
  }
  oc <- h_get_oc_rct(all_sizes = all_sizes, nnr = nnr, Nmax = Nmax, nActive = nActive, nControl = nControl, decision = decision)
  list(
    oc = oc,
    ExpectedN = mean(all_sizes),
    ExpectedNactive = mean(nActive),
    ExpectedNcontrol = mean(nControl),
    Decision = decision,
    SampleSize = all_sizes,
    SampleSizeActive = nActive,
    SampleSizeControl = nControl,
    union_nn = nnr,
    wiggled_nnE = nnrE,
    wiggled_nnF = nnrF,
    wiggle_dist = dist,
    params = as.list(match.call(expand.dots = FALSE))
  )
}
```
# Example of new_RctPostProbDist when `relativeDelta = TRUE` and comparing with vintage function below:
```{r}
set.seed(2000)
new_RctPostprobDist(
  nnE = input$nn,
  pE = input$pE,
  pS = input$pS,
  deltaE = input$deltaE,
  deltaF = input$deltaF,
  relativeDelta = TRUE,
  tL = input$tL,
  tU = input$tU,
  parE = input$parE,
  parS = c(a = 1, b = 1),
  randRatio = 1,
  sim = 50,
  wiggle = FALSE,
  nnF = input$nnF
)
```

```{r}
set.seed(2000)
ocRctPostprobDist(
  nn = input$nn,
  pE = input$pE,
  pS = input$pS,
  deltaE = input$deltaE,
  deltaF = input$deltaF,
  relativeDelta = TRUE,
  tL = input$tL,
  tU = input$tU,
  parE = input$parE,
  parS = input$parS,
  randRatio = 1,
  ns = 50,
  nr = FALSE,
  d = NULL,
  nnF = input$nnF
)
```
# Example of new_ocRctPostprobDist when `relativeDelta = FALSE`.
```{r}
set.seed(2000)
new_RctPostprobDist(
  nnE = input$nn,
  pE = input$pE,
  pS = input$pS,
  deltaE = input$deltaE,
  deltaF = input$deltaF,
  relativeDelta = FALSE,
  tL = input$tL,
  tU = input$tU,
  parE = input$parE,
  parS = input$parE,
  randRatio = 1,
  sim = 50,
  wiggle = FALSE,
  nnF = input$nnF
)
```
# Example of ocRctPostprobDist when `relativeDelta = FALSE`.
```{r}
set.seed(1090)
ocRctPostprobDist(
  nn = input$nn,
  pE = input$pE,
  pS = input$pS,
  deltaE = input$deltaE,
  deltaF = input$deltaF,
  relativeDelta = FALSE,
  tL = input$tL,
  tU = input$tU,
  parE = c(a = 1, b = 1),
  parS = c(a = 1, b = 1),
  randRatio = 1,
  ns = 50,
  nr = FALSE,
  d = NULL,
  nnF = input$nnF
)
```
