---
title: "Design Document for ocPredprob.R"
output: html_document
date: "2024-01-08"
---
design details
multiple looks @ 10, 20, 30 patients
True response rate of the treatment group=0.4

stop for efficacy (deltaE): P(pE > pS + deltaE) >tU
stop for futility (deltaF): P(pE < pS - deltaF) >tL
where pE is the response rate of treatment group, pS is the response rate of
control group. Both of them are random variables.

Design notes :

Three helper functions assist. Two from ocPostprob.R. The only one specialised for ocPostprobDist is `h_get_decisionDist` which gives the end result.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
h_get_decision_predprob <- function(nnr, truep, p0, p1, parE = c(1, 1), nnE, nnF, tL, tU) {
  index_look <- 1
  assert_numeric(nnr)
  size_look <- nnr[index_look]
  all_sizes <- decision <- NA
  response <- stats::rbinom(max(nnr), size = 1, truep)
  assert_numeric(response, lower = 0, upper = 1)
  while (is.na(decision) && index_look <= length(nnr)) {
    if (size_look %in% nnF) {
      qL <- 1 - predprob(x = sum(x[1:i]), n = i, Nmax = Nmax, p = p1, thetaT = 1 - tFu, parE = parE)$result
      assert_number(qL, lower = 0, upper = 1)
      decision <- ifelse(qL >= tL, FALSE, NA)
    }
    if (size_look %in% nnE) {
      qU <- predprob(x = sum(x[1:i]), n = i, Nmax = Nmax, p = p0, thetaT = tT, parE = parE)$result
      assert_number(qU, lower = 0, upper = 1)
      decision <- ifelse(qU < tU, decision, TRUE)
    }
    all_sizes <- size_look
    index_look <- index_look + 1
    size_look <- nnr[index_look]
  }
  list(
    decision = decision,
    all_sizes = all_sizes
  )
}
```

```{r}
h_get_oc_predprob <- function(all_sizes, nnr, decision, nnrE, nnrF) {
  sim <- length(all_sizes)
  assert_logical(decision, len = sim)
  assert_numeric(all_sizes)
  assert_numeric(nnrE, lower = 0, upper = max(nnrE))
  assert_numeric(nnrF, lower = 0, upper = max(nnrF))
  data.frame(
    ExpectedN = mean(all_sizes, na.rm = TRUE),
    PrStopEarly = mean(all_sizes < max(nnrF), na.rm = TRUE),
    PrEarlyEff = sum(decision * (all_sizes < max(nnrE)), na.rm = TRUE) / sim,
    PrEarlyFut = sum((1 - decision) * (all_sizes < max(nnrF)), na.rm = TRUE) / sim,
    PrEfficacy = sum(decision, na.rm = TRUE) / sim,
    PrFutility = sum(1 - decision, na.rm = TRUE) / sim,
  )
}
```


```{r cars}
ocPredprob <- function(nnE,
                       truep,
                       p0,
                       p1,
                       tL,
                       tU,
                       parE = c(1, 1),
                       sim = 50000,
                       wiggle = FALSE,
                       nnF = nnE) {
  nn <- sort(unique(c(nnF, nnE)))
  assert_number(sim, lower = 1, finite = TRUE)
  assert_flag(wiggle)
  if (sim < 50000) {
    warning("Advise to use sim >= 50000 to achieve convergence")
  }
  decision <- vector(length = sim)
  all_sizes <- vector(length = sim)
  for (k in seq_len(sim)) {
    if (length(nn) != 1 && wiggle) {
      dist <- h_get_distance(nn = nn)
      nnr <- h_get_looks(dist = dist, nnE = nnE, nnF = nnF)
      nnrE <- nnr$nnrE
      nnrF <- nnr$nnrF
    } else {
      nnrE <- nnE
      nnrF <- nnF
    }
    nnr <- unique(c(nnrE, nnrF))
    tmp <- h_get_decision_predprob(
      nnr = nnr,
      truep = truep, p0 = p0, p1 = p1,
      parE = c(1, 1), nnE = nnrE,
      nnF = nnrF, tL = tL, tU = tU
    )
    decision[k] <- tmp$decision
    all_sizes[k] <- tmp$all_sizes
  }
  oc <- h_get_oc_predprob(all_sizes = all_sizes, nnr = nnr, decision = decision, nnrE = nnrE, nnrF = nnrF)
  list(
    oc = oc,
    Decision = decision,
    SampleSize = all_sizes,
    union_nn = nnr,
    input_nnE = nnE,
    input_nnF = nnF,
    wiggled_nnE = nnrE,
    wiggled_nnF = nnrF,
    wiggle_dist = dist,
    params = as.list(match.call(expand.dots = FALSE))
  )
}
```

Efficacy Look



```{r pressure, echo=FALSE}
# nnE,
#                        truep,
#                        p0,
#                        p1,
#                        tL,
#                        tU,
#                        parE = c(1, 1),
#                        sim = 50000,
#                        wiggle = FALSE,
#                        nnF = nnE
res2 <- ocPredprob(
  nnE = c(10, 20, 30),
  truep = 0.4,
  p0 = 0.25,
  tT = 0.6,
  phiL = 0.2,
  phiU = 0.8,
  parE = c(1, 1),
  sim = 100,
  wiggle = TRUE
)
res2$oc
```
