---
title: "Design Document for plotOc"
output: html_document
date: "2025-03-05"
editor_options:
  chunk_output_type: console
---
# Purpose 
Plot a bar plot for simulated results of :
- `ocPostprob()` with cases
  1. `length(nnE) = length(nnF)` AND `wiggle = FALSE` ✔️
  2. `length(nnE) = length(nnF)` AND `wiggle = TRUE`. (looks wiggle together)
  3. `length(nnE) =! length(nnF)` AND `wiggle = FALSE`. (looks are different and don't wiggle)
  4. `length(nnE) =! length(nnF)` AND `wiggle = TRUE`. (looks are different and wiggle)
- all other `oc` series such as `ocPredprob()` where `wiggle = TRUE`

Note these results and definitions :
FALSE = met the futility criteria
TRUE = met the efficacy criteria

which can be improved in future PR, e.g "Go" as factor for "met the efficacy criteria". 

## Important packages

```{r}
library(dplyr)
library(tidyr)
```

# `ocPostprob()` 

## 1. `ocPostprob()` where `length(nnE) = length(nnF)` AND `wiggle = FALSE`

```{r}
# Here, we only have one Futility but many Efficacy looks or stop.
res1 <- ocPostprob(
  nnE = c(10, 20, 30), 
  truep = 0.40, 
  p0 = 0.20, 
  p1 = 0.30, 
  tL = 0.60, 
  tU = 0.80, 
  parE = c(1, 1),
  sim = 100, 
  wiggle = FALSE, 
  nnF = c(10, 20, 30)
)

res1 # (may need to add wiggled distance)
res1$oc
```

## 2. `ocPostprob()` where `length(nnE) = length(nnF)` AND `wiggle = TRUE`. (looks wiggle together)

```{r}
# Here, we only have one Futility but many Efficacy looks or stop.
res2 <- ocPostprob(
  nnE = c(10, 20, 30), 
  truep = 0.40, 
  p0 = 0.20, 
  p1 = 0.30, 
  tL = 0.60, 
  tU = 0.80, 
  parE = c(1, 1),
  sim = 100, 
  wiggle = TRUE, 
  nnF = c(10, 20, 30)
)

res2 # (may need to add wiggled distance as param)
res2$oc
```

## 3. `ocPostprob()` where `length(nnE) =! length(nnF)` AND `wiggle = FALSE`. (looks are different and don't wiggle)
```{r}
res3 <- ocPostprob(
  nnE = c(10, 20, 30), 
  truep = 0.40, 
  p0 = 0.20, 
  p1 = 0.30, 
  tL = 0.60, 
  tU = 0.80, 
  parE = c(1, 1),
  sim = 100, 
  wiggle = FALSE, 
  nnF = 30
)

res3 # (may need to add wiggled distance)
res3$oc
```

## 4. `ocPostprob()` where `length(nnE) =! length(nnF)` AND `wiggle = TRUE`. (looks are different and wiggle)

```{r}
res4 <- ocPostprob(
  nnE = c(10, 20, 30), 
  truep = 0.40, 
  p0 = 0.20, 
  p1 = 0.30, 
  tL = 0.60, 
  tU = 0.80, 
  parE = c(1, 1),
  sim = 100, 
  wiggle = TRUE, 
  nnF = 30
)

res4
res4$oc
```

## 5. `ocPredprob()` where decision1 = FALSE
```{r}
res5 <- ocPredprob(
  nnE = c(10, 20),
  truep = 0.6,
  p0 = 0.25,
  p1 = 0.25,
  tT = 0.6,
  tF = 0.6,
  phiU = 0.8,
  phiFu = 0.8,
  parE = c(1, 1),
  sim = 50,
  wiggle = TRUE,
  nnF = c(10, 20),
  decision1 = FALSE
)
res5$oc
```

## 6. `ocPostprobDist()` where wiggle = TRUE
```{r}
res6 <- ocPostprobDist(
  nnE = c(10, 20, 30),
  truep = 0.4,
  deltaE = 0.1,
  deltaF = -0.1,
  tL = 0.6,
  tU = 0.6,
  parE = c(1, 1),
  parS = c(5, 25),
  sim = 50,
  wiggle = TRUE
)
res6$oc
```

## 7. `ocPredprobDist()` where decision1 = TRUE and relativeDelta = FALSE
```{r}
res7 <- ocPredprobDist(
  nnE = c(10, 20, 30),
  truep = 0.40,
  deltaE = 0.10,
  deltaF = 0.10,
  relativeDelta = FALSE,
  tT = 0.6,
  phiU = 0.80,
  phiL = 0.20,
  parE = c(1, 1),
  parS = c(5, 25),
  weights = 1,
  weightsS = 1,
  sim = 50,
  wiggle = TRUE,
  decision1 = TRUE
)
res7$oc
```
## 8. `ocPredprobDist()` where  decision1 = TRUE and relativeDelta = TRUE
```{r}
res8 <- ocPredprobDist(
  nnE = c(10, 20, 30),
  truep = 0.40,
  deltaE = 0.5,
  deltaF = 0.5,
  relativeDelta = TRUE,
  tT = 0.6,
  phiU = 0.80,
  phiFu = 0.7,
  parE = c(1, 1),
  parS = c(5, 25),
  weights = 1,
  weightsS = 1,
  sim = 50,
  nnF = c(10, 20, 30),
  wiggle = TRUE,
  decision1 = TRUE
)
res8
```

## 9. `ocRctPostprobDist()` where relativeDelta = FALSE
```{r}
res9 <- ocRctPostprobDist(
  nnE = c(10, 20, 30),
  pE = 0.4,
  pS = 0.3,
  deltaE = 0.15,
  deltaF = 0.05,
  relativeDelta = FALSE,
  tL = 0.2,
  tU = 0.8,
  parE = c(a = 1, b = 1),
  parS = c(a = 1, b = 1),
  randRatio = 1,
  sim = 50,
  wiggle = TRUE,
  nnF = c(10, 20, 30)
)
res9
```

## 10. `ocRctPostprobDist()` where relativeDelta = FALSE
```{r}
res10 <- ocRctPostprobDist(
  nnE = c(10, 20, 30),
  pE = 0.4,
  pS = 0.3,
  deltaE = 0.15,
  deltaF = 0.05,
  relativeDelta = FALSE,
  tL = 0.2,
  tU = 0.8,
  parE = c(a = 1, b = 1),
  parS = c(a = 1, b = 1),
  randRatio = 1,
  sim = 50,
  wiggle = FALSE,
  nnF = c(10, 20, 30)
)
res10
```

## 11. `ocRctPredprobDist()` where relativeDelta = TRUE
```{r}
res11 <- ocRctPredprobDist(
  nnE = c(10, 20, 30),
  pE = 0.3,
  pS = 0.3,
  deltaE = 0.2,
  deltaF = 0.1,
  phiU = 0.8,
  phiFu = 0.2,
  relativeDelta = TRUE,
  tT = 0.6,
  tF = 0.4,
  parE = c(a = 1, b = 1),
  parS = c(a = 1, b = 1),
  weights = 1,
  weightsS = 1,
  randRatio = 1,
  sim = 50,
  wiggle = TRUE,
  nnF = c(10, 20, 30),
  decision1 = FALSE
)
res11
```

## 12. `ocRctPredtprobDist()` where decision1 = TRUE
```{r}
res12 <- ocRctPredprobDist(
  nnE = c(10, 20, 30),
  pE = 0.3,
  pS = 0.3,
  deltaE = 0.2,
  deltaF = 0.1,
  phiU = 0.8,
  phiFu = 0.2,
  relativeDelta = TRUE,
  tT = 0.6,
  tF = 0.4,
  parE = c(a = 1, b = 1),
  parS = c(a = 1, b = 1),
  weights = 1,
  weightsS = 1,
  randRatio = 1,
  sim = 50,
  wiggle = FALSE,
  nnF = c(10, 20, 30),
  decision1 = TRUE
)
res12
```

### Helper `h_get_dataframe`

```{r}
h_get_dataframe <- function(decision, sample_size, all_looks) {
  df <- data.frame(
    decision = decision,
    sample_size = sample_size,
    look = all_looks
  )
  # summarise into frequency table
  df <- df %>%
    group_by(decision, look) %>%
    summarise(prop = sum(length(decision)) / nrow(df)) %>%
    as_tibble()
  # setting levels of factors
  all_decision <- c(TRUE, FALSE, NA)
  all_looks <- unique(sort(all_looks))
  df$decision <- factor(df$decision, levels = all_decision)
  df$look <- factor(df$look, levels = all_looks)
  df <- df %>%
    complete(decision, look, fill = list(prop = 0))
  df
}
```

# Example for case 1
```{r}
h_get_dataframe(res1$Decision, sample_size = res1$SampleSize, res1$Looks)
```

# Example for case 2
```{r}
h_get_dataframe(res2$Decision, sample_size = res2$SampleSize, res2$Looks)
```

# Example for case 3
```{r}
h_get_dataframe(res3$Decision, sample_size = res3$SampleSize, res3$Looks)
```

# Example for case 4
```{r}
h_get_dataframe(res4$Decision, sample_size = res4$SampleSize, res4$Looks)
```

# Example for case 5  
```{r}
h_get_dataframe(res5$Decision, sample_size = res5$SampleSize, res5$Looks)
```

# Example for case 6 
```{r}
h_get_dataframe(res6$Decision, sample_size = res6$SampleSize, res6$Looks)
```

# Example for case 7 
```{r}
h_get_dataframe(res7$Decision, sample_size = res7$SampleSize, res7$Looks)
```

# Example for case 8 
```{r}
h_get_dataframe(res8$Decision, sample_size = res8$SampleSize, res8$Looks)
```

# Example for case 9 
```{r}
h_get_dataframe(res9$Decision, sample_size = res9$SampleSize, res9$Looks)
```

# Example for case 10 
```{r}
h_get_dataframe(res10$Decision, sample_size = res10$SampleSize, res10$Looks)
```

# Example for case 11 
```{r}
h_get_dataframe(res11$Decision, sample_size = res11$SampleSize, res11$Looks)
```

# Example for case 12 
```{r}
h_get_dataframe(res12$Decision, sample_size = res12$SampleSize, res12$Looks)
```

## `plotOc`

```{r}
plotOc <- function(decision, sample_size, all_looks, wiggle_status) {
  df <- h_get_dataframe(
    decision = decision,
    sample_size = sample_size,
    all_looks = all_looks
  )
  barplot <-
    ggplot2::ggplot(df, ggplot2::aes(fill = decision, x = look, y = prop)) +
    ggplot2::geom_bar(position = "dodge", stat = "identity") +
    ggplot2::ggtitle(
      "Results from simulation : \nProportion of Go/Stop/Grey zone decisions per interim/final analysis"
    ) +
    ggplot2::theme(title = ggplot2::element_text(size = 13)) +
    ggplot2::ylab("percentage") +
    ggplot2::theme(axis.text.x = ggplot2::element_text(size = 12)) +
    ggplot2::xlab("look (n)") +
    ggplot2::scale_fill_manual(
      values = c("#009E73", "#FF0046", "lightgrey"),
      labels = c("Go", "Stop", "Grey zone")
    ) +
    ggplot2::labs(fill = "Decision")
  generic_title <-
    "Results from simulation : \nProportion of Go/Stop/Grey zone decisions per interim/final analysis"
  wiggle_warning_on_title <- paste(
    generic_title, "\nNote that sample sizes may differ slightly from the ones labeled"
  )

  if (wiggle_status) {
    barplot +
      ggplot2::ggtitle(wiggle_warning_on_title)
  } else {
    barplot +
      ggplot2::ggtitle(generic_title)
  }
}
```

### Plot case 1:
```{r}
plotOc(
  decision = res2$Decision,
  sample_size = res2$SampleSize,
  all_looks = res2$Looks,
  wiggle_status = res2$params$wiggle
)
```
### Plot case 2:
```{r}
plotOc(
  decision = res1$Decision,
  sample_size = res1$SampleSize,
  all_looks = res1$Looks,
  wiggle_status = res1$params$wiggle
)
```
### Plot case 3:
```{r}
plotOc(
  decision = res3$Decision,
  sample_size = res3$SampleSize,
  all_looks = res3$Looks,
  wiggle_status = res3$params$wiggle
)
```
### Plot case 4:
```{r}
plotOc(
  decision = res4$Decision,
  sample_size = res4$SampleSize,
  all_looks = res4$Looks,
  wiggle_status = res4$params$wiggle
)
```

### Plot case 5`ocPredprob()`
```{r}
plotOc(
  decision = res5$Decision,
  sample_size = res5$SampleSize,
  all_looks = res5$Looks,
  wiggle_status = res5$params$wiggle
)
```

### Plot case 6 `ocPostProbDist``
```{r}
plotOc(
  decision = res6$Decision,
  sample_size = res6$SampleSize,
  all_looks = res6$Looks,
  wiggle_status = res6$params$wiggle
)
```

### Plot case 7 `ocPredProbDist`
```{r}
plotOc(
  decision = res7$Decision,
  sample_size = res7$SampleSize,
  all_looks = res7$Looks,
  wiggle_status = res7$params$wiggle
)
```

### Plot case 8 `ocPredProbDist`
```{r}
plotOc(
  decision = res8$Decision,
  sample_size = res8$SampleSize,
  all_looks = res8$Looks,
  wiggle_status = res8$params$wiggle
)
```
### Plot case 9`ocRctPostProbDist`
```{r}
plotOc(
  decision = res9$Decision,
  sample_size = res9$SampleSize,
  all_looks = res9$Looks,
  wiggle_status = res9$params$wiggle
)
```
### Plot case 10 `ocRctPostProbDist`
```{r}
plotOc(
  decision = res10$Decision,
  sample_size = res10$SampleSize,
  all_looks = res10$Looks,
  wiggle_status = res10$params$wiggle
)
```
### Plot case 11 `ocRctPredProbDist`
```{r}
plotOc(
  decision = res11$Decision,
  sample_size = res11$SampleSize,
  all_looks = res11$Looks,
  wiggle_status = res11$params$wiggle
)
```
### Plot case 12 `ocRctPredProbDist`
```{r}
plotOc(
  decision = res12$Decision,
  sample_size = res12$SampleSize,
  all_looks = res12$Looks,
  wiggle_status = res12$params$wiggle
)
```
