---
title: "Design Document for plotOc"
output: html_document
date: "2025-03-05"
editor_options:
  chunk_output_type: console
---
# Purpose 
Plot a bar plot for simulated results of :
- `ocPostprob()` with cases
  1. `length(nnE) = length(nnF)` AND `wiggle = FALSE` ✔️
  2. `length(nnE) = length(nnF)` AND `wiggle = TRUE`. (looks wiggle together)
  3. `length(nnE) =! length(nnF)` AND `wiggle = FALSE`. (looks are different and don't wiggle)
  4. `length(nnE) =! length(nnF)` AND `wiggle = TRUE`. (looks are different and wiggle)
- `ocPredprob()`
- others

Note these results and definitions :
FALSE = met the futility criteria
TRUE = met the efficacy criteria

which can be improved in future PR, e.g "Go" as factor for "met the efficacy criteria". 

## Important packages

```{r}
library(dplyr)
```

### Input parameters :

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

input <- list(
  nn = c(10, 20, 30),
  nnE = c(10, 20, 30),
  nnF = c(10, 20, 30),
  p0 = 0.2,
  p1 = 0.3,
  parE = c(1, 1),
  truep = 0.4,
  tU = 0.8,
  tL = 0.6,
  wiggle = TRUE,
  sim = 100
)
```

# `ocPostprob()` 

## 1. `ocPostprob()` where `length(nnE) = length(nnF)` AND `wiggle = FALSE`

```{r}
# Here, we only have one Futility but many Efficacy looks or stop.
res <- ocPostprob(
  nnE = c(10, 20, 30), truep = 0.40, p0 = 0.20, p1 = 0.30, tL = 0.60, tU = 0.80, parE = c(1, 1),
  sim = 100, wiggle = FALSE, nnF = c(10, 20, 30)
)

res # (may need to add wiggled distance)
res$oc
```

## 2. `ocPostprob()` where `length(nnE) = length(nnF)` AND `wiggle = TRUE`. (looks wiggle together)

```{r}
# Here, we only have one Futility but many Efficacy looks or stop.
res <- ocPostprob(
  nnE = c(10, 20, 30), truep = 0.40, p0 = 0.20, p1 = 0.30, tL = 0.60, tU = 0.80, parE = c(1, 1),
  sim = 100, wiggle = TRUE, nnF = c(10, 20, 30)
)

res # (may need to add wiggled distance as param)
res$oc
```

## 3. `ocPostprob()` where `length(nnE) =! length(nnF)` AND `wiggle = FALSE`. (looks are different and don't wiggle)
```{r}
res <- ocPostprob(
  nnE = c(10, 20, 30), truep = 0.40, p0 = 0.20, p1 = 0.30, tL = 0.60, tU = 0.80, parE = c(1, 1),
  sim = 100, wiggle = FALSE, nnF = 30
)

res # (may need to add wiggled distance)
res$oc
```

## 4. `ocPostprob()` where `length(nnE) =! length(nnF)` AND `wiggle = TRUE`. (looks are different and wiggle)

```{r}
res <- ocPostprob(
  nnE = c(10, 20, 30), truep = 0.40, p0 = 0.20, p1 = 0.30, tL = 0.60, tU = 0.80, parE = c(1, 1),
  sim = 100, wiggle = TRUE, nnF = 30
)

res # (may need to add wiggled distance)
res$oc
```


### Curating data table such that it can be an input to `geom_bar()` for all cases

```{r}

h_get_dataframe <- function(decision, sample_size, all_looks) {
  df <- data.frame(decision = decision,
                      sample_size = sample_size,
                      look = all_looks)
  # summarise into frequency table
  df <- df %>%  
  group_by(decision, sample_size) %>% 
    summarise(prop = sum(length(decision))/nrow(df)) %>% 
    as_tibble()
  # setting levels of factors
  all_decision <- c(TRUE, FALSE, NA)
  all_looks <- c(10, 20, 30)
  df$decision <- factor(df$decision, levels = all_decision)
  df$sample_size <- factor(df$sample_size, levels = all_looks)
  # data input for plot, such that all variables present, even empty cases
  df <- df %>%
    complete(decision, nesting(sample_size), 
             fill = list(prop = 0))
  ggplot(df, aes(fill = decision, x = sample_size, y = prop)) +
      geom_bar(position = "dodge", stat = "identity")
}

h_get_dataframe(res$Decision, sample_size = res$SampleSize, res$Looks)

# create df from res
col_data = data.frame(decision = res$Decision,
                      sample_size = res$SampleSize,
                      look = res$Looks)
# check data summary
data <- col_data %>% 
  group_by(decision, sample_size) %>% 
  summarise(prop = sum(length(decision))/nrow(col_data)) %>% as_tibble()
data

# setting levels of factors
all_decision <- c(TRUE, FALSE, NA)
all_looks <- c(10, 20, 30)
data$decision <- factor(data$decision, levels = all_decision)
data$sample_size <- factor(data$sample_size, levels = all_looks)

# data input for plot, such that all variables present, even empty cases
data1 <- data %>%
  complete(decision, nesting(sample_size), 
           fill = list(prop = 0))

ggplot(data1, aes(fill = decision, x = sample_size, y = prop)) +
  geom_bar(position = "dodge", stat = "identity")
```

## Testing four cases

```{r}

# h_ ----
test_that("h_get_decision will give GO decision in favourable conditions", {
  set.seed(1989)
  result <- h_get_decision(
  expect_equal(result$decision, TRUE)
  expect_equal(result$all_sizes, 20)
})

```


## User facing function

```{r}
plotOc <- function(oc) {

  if (wiggle == FALSE) {
    data <-  table(oc$Decision, oc$SampleSize) / oc$params$sim
  } else {

  }

ggplot(oc, aes(x = name, y = value)) +
  geom_bar(stat = "identity") +
  ggtitle("Percentage of trials that Go and Stop per look") +
  ylabs("Percentage %") +
  xlabs("Looks and sample size")

  ## plot function for oc.predprob or oc.postprob, or the dist versions of them
  graphics::barplot(table(oc$Decision, oc$SampleSize) / oc$params$sim, beside = TRUE)


  ## get the parameter
  parDat <- lapply(z$params, deparse)

  ## get parameters, which are saved in parDat
  allParNames <- c(
    "p", "p0", "p1",
    "delta", "deltaE", "deltaF", "relativeDelta",
    "phiL", "phiU", "tL", "tU", "tT", "parE", "parS"
  )
  parInds <- which(names(parDat) %in% allParNames)

  graphics::title(paste(names(parDat[parInds]), parDat[parInds],
    sep = "=", collapse = ", "
  ))
  graphics::title(xlab = paste(dimnames(z$oc)[[2]], signif(z$oc, 3),
    sep = " = ", collapse = ", "
  ))

  return(invisible())
}
```
